---
title: "Wrangling Senegal Census data with tidyverse tools"
author: "IBRAHIM KASSOUM Habibou & HEMA Aboubacar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
```


## Project Sénégal

### Preambule: 

```{r}
rm(list=ls())
```


```{r Package needed, results = "hide"}


## Importing library
### List of required packages
required_packages <- c("tidyverse","janitor" ,"readr","dplyr","haven","sf",
                       "flextable","sp", "factoextra", "FactoMineR","gtsummary", "sjPlot")

# Check if packages are installed
missing_packages <- setdiff(required_packages, installed.packages()[,"Package"])

### Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

### Load all packages
lapply(required_packages, library, character.only = TRUE)

```


### Importing the dataset

```{r}

# Read shapefile data for 2002 and 2013
sp_rgph_2002 <- sf::read_sf(paste0(here::here(),"/data/SHAPEFILE/EAs_2002.shp"))

sp_rgph_2013 <- sf::read_sf(paste0(here::here(),"/data/SHAPEFILE/EAs_2013.shp"))

# Display the first 10 rows of the 2002 and 2013 shapefiles
head(sp_rgph_2002, 10L)
head(sp_rgph_2013, 10L)
```
### Importing the dataset an checking for duplicated row

```{r}
# Read and check for duplicated rows in 2002 datasets

# Read and check for duplicated rows in the "deces_2002" dataset
deces_2002 <- read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 Deces.sav"))
deces_2002 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "deces_complement" dataset
deces_2002_complement <- read_sav(paste0(here::here(),"/data/1. RGPH 2002/PDBA_rgph2002_Deces Complement Pikine.sav"))
deces_2002_complement %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "individus_2002" dataset
individus_2002 <- haven::read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 Individus.v2.sav"))
individus_2002 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "menage_2002" dataset
menage_2002 <- haven::read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 menage.sav"))
menage_2002 %>% janitor::get_dupes()
menage_2002 %>% janitor::get_dupes(ID_MENAGE)

# Read and check for duplicated rows in the "menage_2002_complement" dataset
menage_2002_complement <- read_sav(paste0(here::here(),"/data/1. RGPH 2002/PDBA_rgph2002_menage_complement.sav"))
menage_2002_complement %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "individus_2002_complement_pikine" dataset
individus_2002_complement_pikine <- read_sav(paste0(here::here(),"/data/1. RGPH 2002/PDBA_rgph2002_Individus Complement Pikine.sav"))
individus_2002_complement_pikine %>% janitor::get_dupes() 

# Joining with the 2002 dataset for deceased individuals
deces_2002 <- deces_2002_complement %>% 
  mutate(ID_MENAGE = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE, QA08, QA09, QA10),
         DR = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE)) %>%
  plyr::rbind.fill(deces_2002)

# Joining with the 2002 dataset for individual data
individus_2002 <- individus_2002_complement_pikine %>% 
  mutate(ID_MENAGE = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE, QA08, QA09, QA10),
         DR = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE)) %>%
  plyr::rbind.fill(individus_2002)

# Joining with the 2002 dataset for household data
menage_2002 <- menage_2002_complement %>% 
  mutate(ID_MENAGE = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE, QA08, QA09, QA10),
         DR = paste0(QA01, QA02, QA03, QA04, QA05, QA06, "0", QA07, QA072_CLE)) %>%
  plyr::rbind.fill(menage_2002)
```


```{r}

# Read and check for duplicated rows in 2013 datasets

# Read and check for duplicated rows in the "deces_2013" dataset
deces_2013 <- read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 Deces_Guediawaye.sav"))
deces_2013 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "individus_2013" dataset
individus_2013 <- haven::read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 Indidivus GUEDIAWAYE.sav"))
individus_2013 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "menage_2013" dataset
menage_2013 <- haven::read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 menage.sav"))
menage_2013 %>% janitor::get_dupes() 
menage_2013 %>% janitor::get_dupes(ID_MENAGE)
```

```{r}
# Creating a variable "RGPH" to identify each survey 

# For 2002
## Add a new column "RGPH" with the value "2002" to the "individus_2002" dataset
individus_2002 <- 
  individus_2002 %>% 
  mutate(RGPH = "2002")

# For 2013
## Add a new column "RGPH" with the value "2013" to the "individus_2013" dataset
individus_2013 <- 
  individus_2013 %>% 
  mutate(RGPH = "2013")
```


### -	Calcul des indices sur les variables composants liés au ménage (Age des chefs de ménage, par sexe entre les deux recensements

```{r}
# Select the CM (Chef de ménage) and compute the average age

# Select relevant columns (B04, B06, B08, RGPH) from the "individus_2013" dataset
individus_2013 %>% 
  select(B04, B06, B08, RGPH) %>%

  # Filter to select only records where "B04" (Chef de ménage) equals 1
  # This is done to identify the household head
  filter(B04== 1) %>% # Household head (B04 == 1) "Chef de ménage"

  # Rename columns for clarity
  rename(
    sexe_cm = B06,        # Renamed B06 to "sexe_cm"
    lien_cm = B04,        # Renamed B04 to "lien_cm"
    age_cm = B08,         # Renamed B08 to "age_cm"
  ) %>% 

  # Combine this filtered data with a similar operation on the "individus_2002" dataset
  bind_rows(
    individus_2002 %>% 
      select(QB03, QB04, QB06, RGPH) %>%
    
    # Filter to select only records where "QB03" (Chef de ménage) equals "Chef de menage"
    filter(QB03 == 1 ) %>% # Household head (QB03 == 1) "Chef de menage"
      
    # Rename columns for consistency
    rename(
      sexe_cm = QB04,        # Renamed QB04 to "sexe_cm"
      lien_cm = QB03,        # Renamed QB03 to "lien_cm"
      age_cm = QB06,         # Renamed QB06 to "age_cm"
    ) 
    
  ) %>% 

  # Remove the "lien_cm" (link to household head) column
  select(-lien_cm) %>% 
  mutate( sexe_cm = ifelse(sexe_cm ==2, 0, 1)) %>% 
  
  # Generate a summary table using "tbl_summary" for the specified columns
  tbl_summary(
    by = RGPH,  # Group by "RGPH" column
    label = list(sexe_cm ~ "Sexe du chef de ménage (Masculin = 1, Féminin = 0)",
                 age_cm ~ "Age du chef de ménage (en année)"),
    statistic = list(all_continuous() ~ "{mean}" 
                     #all_categorical() ~ "{p}"
                     ),
    digits = everything() ~ c(0,0),
    missing = "no"
  ) %>% 

  # Modify the table header to provide a descriptive label
  modify_header(label ~ "Caractéristique du Chef de ménage entre les deux recensements") %>%
  
  # Add difference statistics to the table
  add_difference() 
```



### -	Calcul multivarié pour des indices de pauvreté (bien des ménages, Niveau d’instruction, accès aux services (bien d’aisance, eau, etc.)

Nettoyage de la variable education
```{r}

# Transform values in the "QB15" column of the "individus_2002" dataset
# The transformation categorizes educational levels into descriptive labels

individus_2002 <- individus_2002 %>% 
  
  mutate(QB15 = case_when(
    
    QB15 == 0 ~ "aucune",      # If QB15 is 0, set it to "aucune"
    
    QB15 == 1 ~ "prescolaire", # If QB15 is 1, set it to "prescolaire"
    
    QB15 %in% seq(2,7,1) ~ "elementaire", # If QB15 is 2 to 7, set it to "elementaire"
    
    QB15 %in% seq(8,11,1) ~ "moyen", # If QB15 is 8 to 11, set it to "moyen"
    
    QB15 %in% seq(12,14,1) ~ "secondaire", # If QB15 is 12 to 14, set it to "secondaire"
    
    QB15 %in% seq(15,19,1) ~ "superieur", # If QB15 is 15 to 19, set it to "superieur"
    
    TRUE ~ as.character(QB15) # If none, keep the old value
    ) %>% 
  structure(label = Hmisc::label(individus_2002$QB15))) # Apply labels to the QB15 column


# Transform values in the "B32" column of the "individus_2013" dataset
# The transformation categorizes educational levels from codes into descriptive labels

individus_2013 <- individus_2013 %>% 
  
  mutate(B32 = case_when(
    
    B32 %in% seq(1,3,1)  ~ "prescolaire", # If B32 is in pre-school levels, set it to "prescolaire" for c("Petite section","Moyenne section", "Grnde section")
    
    B32 %in% seq(4,9,1) ~ "elementaire", # If B32 corresponds to elementary levels, set it to "elementaire" for c("Cours d'initiation (CI)", "Cours préparatoire (CP)", "Cours élémentaire 1ère année (CE1)", "Cours élémentaire 2ème année (CE2)", "Cours moyen 1ère année (CM1)", "Cours moyen 2ème année (CM2)"
    
    B32 %in% seq(10,13,1) ~ "moyen", # If B32 corresponds to middle school levels, set it to "moyen" for c("Sixième (6ième)", "Cinquième (5ième)", "Quatrième (4ième)", "Troisième (3ième)")
    
    B32 %in% seq(14,16,1) ~ "secondaire", # If B32 corresponds to high school levels, set it to "secondaire" for c("Seconde", "Première", "Terminale")
    
    B32 %in% seq(17,24,1) ~ "superieur", # If B32 corresponds to higher education levels, set it to "superieur" c("1ère année", "2ème année", "Licence/3ième année", "Maitrise/4ième année", "DEA/5ième année", "6ième année", "7ième année", "8ième année et +")
    TRUE ~ as.character(B32) %>% 
  structure(label = Hmisc::label(individus_2013$B32)))) # Apply labels to the B32 column
```

Education
SDG 4: Quality of Education (Year of Schooling)

```{r}

# For the 2002 individus data  

individus_2002 <- 
  individus_2002 %>% 
  left_join(
# Group the 'individus_2002' dataframe by the 'ID_MENAGE' column
individus_2002 %>%
  group_by(ID_MENAGE) %>%
  
# Calculate the sum of individuals with 'elementaire' education within each household
  summarise(nbr_educ_primaire = sum(QB15 == "elementaire", na.rm = TRUE),
  
# Calculate the sum of individuals with 'moyen' education within each household
            nbr_educ_moyen = sum(QB15 == "moyen", na.rm = TRUE),
            
# Calculate the sum of individuals with 'secondaire' education within each household
            nbr_educ_secondaire = sum(QB15 == "secondaire", na.rm = TRUE),
            
# Calculate the sum of individuals with 'superieur' education within each household
            nbr_educ_superieur = sum(QB15 == "superieur", na.rm = TRUE)
  ) %>%
  
# Remove the grouping to get a flat summary data frame
  ungroup() %>%
  
# Calculate a binary variable 'school_attendance' for each household.
# If the sum of education levels is greater than 0, it means at least one individual in the household is attending school (0), 
# otherwise, it means no one is attending school (1).
  mutate(years_schooling_dim = 
           as.factor(ifelse(nbr_educ_primaire + nbr_educ_moyen + nbr_educ_secondaire + nbr_educ_superieur > 0, 0, 1)) %>% structure(label = "Education: years of schooling")) %>%
  select(ID_MENAGE, years_schooling_dim),

  by=c("ID_MENAGE"="ID_MENAGE")
)
  
# For the 2013 individus data  

individus_2013 <- 
  individus_2013 %>% 
  left_join(
# Group the 'individus_2013' dataframe by the 'ID_MENAGE' column
individus_2013 %>%
  group_by(ID_MENAGE) %>%
  
# Calculate the sum of individuals with 'elementaire' education within each household
  summarise(nbr_educ_primaire = sum(B32 == "elementaire", na.rm = TRUE),
  
# Calculate the sum of individuals with 'moyen' education within each household
            nbr_educ_moyen = sum(B32 == "moyen", na.rm = TRUE),
            
# Calculate the sum of individuals with 'secondaire' education within each household
            nbr_educ_secondaire = sum(B32 == "secondaire", na.rm = TRUE),
            
# Calculate the sum of individuals with 'superieur' education within each household
            nbr_educ_superieur = sum(B32 == "superieur", na.rm = TRUE)
  ) %>%
  
# Remove the grouping to get a flat summary data frame
  ungroup() %>%
  
# Calculate a binary variable 'school_attendance' for each household.
# If the sum of education levels is greater than 0, it means at least one individual in the household is attending school (0), 
# otherwise, it means no one is attending school (1).
  mutate(years_schooling_dim = 
           as.factor(ifelse(nbr_educ_primaire + nbr_educ_moyen + nbr_educ_secondaire + nbr_educ_superieur > 0, 0, 1)) %>% structure(label = "Education: years of schooling")) %>%
  select(ID_MENAGE, years_schooling_dim),

  by=c("ID_MENAGE"="ID_MENAGE")
)
```

SDG 4: School Attendance (Year of Schooling)

```{r}

# Processing for the year 2002
individus_2002 <- 
  individus_2002 %>% 
  left_join(
    # Joining the data with summarized statistics per household
    individus_2002 %>%
    group_by(ID_MENAGE) %>%
    summarise(
      # Calculating the number of children (lien_cm=="children" and age < 13) with no school attendance 
      
      nbr_school_attendance = sum(QB15 == "aucune" & age5 > 3 & QB03 == 3, na.rm = TRUE),
      
      #The total number of children in the household
      nbr_children_hh = sum(QB03 == 3)
    ) %>%
    ungroup() %>%
      
    # Computing a dimension for school attendance in the household
    mutate(
      school_attendance_dim = as.factor(ifelse(nbr_school_attendance == nbr_children_hh, 1, 0))%>% structure(label = "Education: school attendance")
    ) %>%
      
    # Selecting relevant columns for the join
    select(ID_MENAGE, school_attendance_dim),
    by = c("ID_MENAGE"="ID_MENAGE")
  )

# Processing for the year 2013
individus_2013 <- 
  individus_2013 %>% 
  left_join(
    # Joining the data with summarized statistics per household
    individus_2013 %>%
    group_by(ID_MENAGE) %>%
    summarise(
      
      # Calculating the number of children (lien_cm=="children" and age < 13) with no school attendance  (NA for B32)
      nbr_school_attendance = sum(is.na(B32) & B08 > 13 & B04 == 3, na.rm = TRUE),
      
      # Computing a dimension for school attendance in the household
      nbr_children_hh = sum(B04 == 3)
    ) %>%
    ungroup() %>%
      
    # Computing a dimension for school attendance in the household
    mutate(
      school_attendance_dim = as.factor(ifelse(nbr_school_attendance == nbr_children_hh, 1, 0)) %>% structure(label = "Education: school attendance")
    ) %>%
      
    # Selecting relevant columns for the join
    select(ID_MENAGE, school_attendance_dim),
    by = c("ID_MENAGE"="ID_MENAGE")
  )
```


Standard of Living

```{r}

# Renaming variables so that they can have the same name in the two
# datasets

# Renaming variables so that they can have the same name in the two datasets

## For the RGPH 2002
menage_2002 <- menage_2002 %>% 
  
  # Mutate and rename variables in the "menage_2002" dataset
  mutate(
    
    nb_indiv = E01 %>% structure(label=Hmisc::label(menage_2002$Nb_indiv)),
    id_dr = DR %>% structure(label=Hmisc::label(menage_2002$DR)),
    
    type_log = as.factor(E01) %>% structure(label=Hmisc::label(menage_2002$E01)),
    
    nbr_piece_occup = as.factor(E02) %>% structure(label=Hmisc::label(menage_2002$E02)),
    
    stat_occupation = as.factor(E03) %>% structure(label=Hmisc::label(menage_2002$E03)),
    
    type_aisance = as.factor(E04) %>% structure(label=Hmisc::label(menage_2002$E04)),
    
    approv_eau = as.factor(E05) %>% structure(label=Hmisc::label(menage_2002$E05)),
    
    eclairage = as.factor(E06) %>% structure(label=Hmisc::label(menage_2002$E06)),
    
    combustible_cuisine = as.factor(E07) %>% structure(label=Hmisc::label(menage_2002$E07)),
    
    nature_mur = as.factor(E08) %>% structure(label=Hmisc::label(menage_2002$E08)),
    nature_toit = as.factor(E09) %>% structure(label=Hmisc::label(menage_2002$E09)),
    nature_sol = as.factor(E10) %>% structure(label=Hmisc::label(menage_2002$E10)),
    radio = as.factor(E11B) %>% structure(label=Hmisc::label(menage_2002$E11B)),
    televiseur = as.factor(E11C) %>% structure(label=Hmisc::label(menage_2002$E11C)),
    video = as.factor(E11D) %>% structure(label=Hmisc::label(menage_2002$E11D)),
    refri_conge = as.factor(E11E) %>% structure(label=Hmisc::label(menage_2002$E11E)),
    
    telephone = as.factor(E11F) %>% structure(label=Hmisc::label(menage_2002$E11F)),
    rechaud_cuis = as.factor(E11G) %>% structure(label=Hmisc::label(menage_2002$E11G)),
    foyer_ameli = as.factor(E11H) %>% structure(label=Hmisc::label(menage_2002$E11H)),
    climatisateur = as.factor(E11I) %>% structure(label=Hmisc::label(menage_2002$E11I)),
    machine_coudre = as.factor(E11J) %>% structure(label=Hmisc::label(menage_2002$E11J)),
    #aucun_11K = E11K %>% structure(label=Hmisc::label(menage_2002$E11K)),
    voiture = as.factor(E12B) %>% structure(label=Hmisc::label(menage_2002$E12B)),
    
    mobylette = as.factor(E12C) %>% structure(label=Hmisc::label(menage_2002$E12C)),
    bicyclette = as.factor(E12E) %>% structure(label=Hmisc::label(menage_2002$E12E)),
    
    caleche_charette = as.factor(E12F) %>% structure(label=Hmisc::label(menage_2002$E12F)),
    
    pirogue = as.factor(E12G) %>% structure(label=Hmisc::label(menage_2002$E12G)),
    
    
    houe_charrue = as.factor(E13B) %>% structure(label=Hmisc::label(menage_2002$E13B)),
    
    caleche_char = as.factor(E13C) %>% structure(label=Hmisc::label(menage_2002$E13C)),
    
    animaux_traite = as.factor(E13D) %>% structure(label=Hmisc::label(menage_2002$E13D)),
                               
    tracteur = as.factor(E13E) %>% structure(label=Hmisc::label(menage_2002$E13E)),
    voiture_cam = as.factor(E13F) %>% structure(label=Hmisc::label(menage_2002$E13F)),
    mobylette_bicycl = as.factor(E13G) %>% structure(label=Hmisc::label(menage_2002$E13G)),
    
    refriger_congelat = as.factor(E13I) %>% structure(label=Hmisc::label(menage_2002$E13I)),
    
    machine_coudre = as.factor(E13J) %>% structure(label=Hmisc::label(menage_2002$E13J)),
    
    materiel_musique = as.factor(E13K) %>% structure(label=Hmisc::label(menage_2002$E13K)),
    
    chaise_bache = as.factor(E13L) %>% structure(label=Hmisc::label(menage_2002$E13L)),
    
    telephone_tele = as.factor(E13M) %>% structure(label=Hmisc::label(menage_2002$E13M)),
    
    photocopieuse = as.factor(E13N) %>% structure(label=Hmisc::label(menage_2002$E13N)),
    
    ordinateur = as.factor(E13O) %>% structure(label=Hmisc::label(menage_2002$E13O)),
    moulin = as.factor(E13P) %>% structure(label=Hmisc::label(menage_2002$E13P)),
    appareil_photo = as.factor(E13Q) %>% structure(label=Hmisc::label(menage_2002$E13Q)),
    
    terrain_bat = as.factor(E13R) %>% structure(label=Hmisc::label(menage_2002$E13R)),
    
    #aucun_produc = E13S %>% structure(label=Hmisc::label(menage_2002$E13S)),
    evacut_ordures = as.factor(E14) %>% structure(label=Hmisc::label(menage_2002$E14)),
    
    evacut_eaux_usees = as.factor(E15) %>% structure(label=Hmisc::label(menage_2002$E15)),
    
    repas_saute = as.factor(E16) %>% structure(label=Hmisc::label(menage_2002$E16)),
    soins_medicaux = as.factor(E17) %>% structure(label=Hmisc::label(menage_2002$E17))
  )


# Mutate and rename variables in the "menage_2013" dataset
menage_2013 <- menage_2013 %>% 
  
  mutate(
    id_dr = IDDR %>% structure(label=Hmisc::label(menage_2013$IDDR)),
    nb_indiv = Nbre_indiv %>% structure(label=Hmisc::label(menage_2013$Nbre_indiv)),
    
    type_log = as.factor(E01) %>% structure(label=Hmisc::label(menage_2013$E01)),
    nbr_piece_occup = as.factor(E02) %>% structure(label=Hmisc::label(menage_2013$E02)),
    
    stat_occupation = as.factor(E03) %>% structure(label=Hmisc::label(menage_2013$E03)),
    
    type_aisance = as.factor(E08) %>% structure(label=Hmisc::label(menage_2013$E08)),
    approv_eau = as.factor(E09) %>% structure(label=Hmisc::label(menage_2013$E09)),
    eclairage = as.factor(E11) %>% structure(label=Hmisc::label(menage_2013$E11)),
    combustible_cuisine = as.factor(E12) %>% structure(label=Hmisc::label(menage_2013$E12)),
    
    nature_mur = as.factor(E05) %>% structure(label=Hmisc::label(menage_2013$E05)),
    nature_toit = as.factor(E06) %>% structure(label=Hmisc::label(menage_2013$E06)),
    nature_sol = as.factor(E07) %>% structure(label=Hmisc::label(menage_2013$E07)),
    radio = as.factor(E13_1) %>% structure(label=Hmisc::label(menage_2013$E13_1)),
    televiseur = as.factor(E13_2) %>% structure(label=Hmisc::label(menage_2013$E13_2)),
    
    video = as.factor(E13_3) %>% structure(label=Hmisc::label(menage_2013$E13_3)),
    refri_conge = as.factor(E13_4) %>% structure(label=Hmisc::label(menage_2013$E13_4)),
    
    telephone = as.factor(E13_6) %>% structure(label=Hmisc::label(menage_2013$E13_6)),
    rechaud_cuis = as.factor(E13_18) %>% structure(label=Hmisc::label(menage_2013$E13_18)),
    
    foyer_ameli = as.factor(E13_7) %>% structure(label=Hmisc::label(menage_2013$E13_7)),
    
    climatisateur = as.factor(E13_8) %>% structure(label=Hmisc::label(menage_2013$E13_8)),
    
    machine_coudre = as.factor(E13_9) %>% structure(label=Hmisc::label(menage_2013$E13_9)),
    
    voiture = as.factor(E15_5) %>% structure(label=Hmisc::label(menage_2013$E15_5)),
    mobylette = as.factor(E15_6) %>% structure(label=Hmisc::label(menage_2013$E15_6)),
    bicyclette = as.factor(E14_3) %>% structure(label=Hmisc::label(menage_2013$E14_3)),
    
    caleche_charette = as.factor(E14_4) %>% structure(label=Hmisc::label(menage_2013$E14_4)),
    
    pirogue = as.factor(E15_7) %>% structure(label=Hmisc::label(menage_2013$E15_7)),
    houe_charrue = as.factor(E15_1) %>% structure(label=Hmisc::label(menage_2013$E15_1)),
    
    caleche_char = as.factor(E15_2) %>% structure(label=Hmisc::label(menage_2013$E15_2)),
    
    animaux_traite = as.factor(E15_3) %>% structure(label=Hmisc::label(menage_2013$E15_3)),
    
    tracteur = as.factor(E15_4) %>% structure(label=Hmisc::label(menage_2013$E15_4)),
    voiture_cam = as.factor(E15_5) %>% structure(label=Hmisc::label(menage_2013$E15_5)),
    
    mobylette_bicycl = as.factor(E15_6) %>% structure(label=Hmisc::label(menage_2013$E15_6)),
    
    refriger_congelat = as.factor(E15_8) %>% structure(label=Hmisc::label(menage_2013$E15_8)),
    
    machine_coudre = as.factor(E15_9) %>% structure(label=Hmisc::label(menage_2013$E15_9)),
    
    materiel_musique = as.factor(E15_10) %>% structure(label=Hmisc::label(menage_2013$E15_10)),
    
    chaise_bache = as.factor(E15_11) %>% structure(label=Hmisc::label(menage_2013$E15_11)),
    
    telephone_tele = as.factor(E15_12) %>% structure(label=Hmisc::label(menage_2013$E15_12)),
    
    photocopieuse = as.factor(E15_13) %>% structure(label=Hmisc::label(menage_2013$E15_13)),
    
    ordinateur = as.factor(E15_14) %>% structure(label=Hmisc::label(menage_2013$E15_14)),
    
    moulin = as.factor(E15_15) %>% structure(label=Hmisc::label(menage_2013$E15_15)),
    appareil_photo = as.factor(E15_16) %>% structure(label=Hmisc::label(menage_2013$E15_16)),
    
    terrain_bat = as.factor(E15_17) %>% structure(label=Hmisc::label(menage_2013$E15_17)),
    
    evacut_ordures = as.factor(E16) %>% structure(label=Hmisc::label(menage_2013$E16)),
    
    evacut_eaux_usees = as.factor(E17) %>% structure(label=Hmisc::label(menage_2013$E17)),
    
    repas_saute = as.factor(E18) %>% structure(label=Hmisc::label(menage_2013$E18)),
    soins_medicaux = as.factor(E20) %>% structure(label=Hmisc::label(menage_2013$E20))
    
    )



```

Données source: 

* Pour les sanitaires: [link](https://www.who.int/data/gho/data/indicators/indicator-details/GHO/population-using-improved-sanitation-facilities-(-))

* Pour l'eau:
[link](https://data.unicef.org/topic/water-and-sanitation/drinking-water/)

```{r}

# For RGPH 2002
menage_2002 <- menage_2002 %>% 
  # Creating a dimension for cooking fuel, checking if it's wood or charcoal
  # Bois=1, Charbon = 2 and Autre=5
  mutate(cooking_fuel_dim = as.factor(ifelse(combustible_cuisine == 1 | combustible_cuisine == 2| combustible_cuisine == 5, 1, 0)) %>% structure(label = "Standard of living: cooking fuel"), 
         
  # Creating a dimension for sanitation, checking for specific types (public edifice, nature, other)
  # Edicule public==4, Dans la nature==4, Autre==6
         sanitation_dim = as.factor(ifelse(type_aisance %in% seq(4, 6), 1, 0)) %>% structure(label = "Standard of living: sanitation"),
         
  # Creating a dimension for drinking water, considering specific sources (indoor well, outdoor well, watercourse, other)
  # Puits interieur = 1, Puits exterieur = 2, cours d'eau = 7 et autre = 8
         drinking_water_dim = as.factor(ifelse(approv_eau %in% c(1, 2, 7, 8), 1, 0)) %>% structure(label = "Standard of living: drinking water"),
         
  # Creating a dimension for electricity, considering different sources (storm lamp, candle, wood, other)
  # Lampe tempête = 6, Bougie = 7, Bois=8, Autre= 9,
         electricity_dim = as.factor(ifelse(eclairage %in% c(6, 7, 8, 9), 1, 0))%>% structure(label = "Standard of living: electricity"),
         
  # Creating a dimension for housing quality, checking specific types of construction materials for walls, floor, and roof
        # For nature_mur, 3 is "Banco", 4 is "Paille/tige" and 5 is 'autre".
         # For nature_sol, 3 is 'Argile/banco", 4 is Sable and 5 is "autre".
         # For nature_toit, 4 is "Chaume" and 5 is "autre"
         housing_dim = as.factor(ifelse(nature_mur %in% c(3, 4, 5) & nature_sol %in% c(3, 4, 5) & nature_toit %in% c(4, 5), 1, 0))%>% structure(label = "Standard of living: housing"),
         
  # Creating a dimension for household assets, checking if various assets are absent in the household
  
         assets_dim = as.factor(ifelse(radio == 0 & televiseur == 0 & telephone == 0 & animaux_traite == 0 & bicyclette == 0 & mobylette == 0 & mobylette_bicycl == 0 & 
                           refri_conge == 0 & refriger_congelat == 0 & voiture == 0 & voiture_cam == 0, 1, 0)) %>% structure(label = "Standard of living: assets")
  )



# For RGPH 2013
menage_2013 <- menage_2013 %>% 
  # Creating a dimension for cooking fuel, checking if it's wood or charcoal
  # Bois=1, Charbon = 2, Bouse de vache=5 and Autre=8
  mutate(cooking_fuel_dim = as.factor(ifelse(combustible_cuisine %in% c(1,2,5,8), 1, 0))%>% structure(label = "Standard of living: cooking fuel"), 
         
  # Creating a dimension for sanitation, checking for specific types (public edifice, nature, other)
  # Non couverte =22, Edicule public==31, Dans la nature==41, Autre==88
         sanitation_dim = as.factor(ifelse(type_aisance %in% c(22, 31, 41, 88), 1, 0))%>% structure(label = "Standard of living: sanitation"),
         
  # Creating a dimension for drinking water, considering specific sources (indoor well, outdoor well, watercourse, other)
  # Puits non protégés = 6, Eau de pluie = 9, Source, cour d'eau = 10 and autre = 98
         drinking_water_dim = as.factor(ifelse(approv_eau %in% c(6, 9, 10, 98), 1, 0))%>% structure(label = "Standard of living: drinking water"),
         
  # Creating a dimension for electricity, considering different sources (storm lamp, candle, wood, other)
  # Lampe à pétrole artisanale=5, Lampe tempête = 6, Bougie = 7, Bois=8, Lampe rechargeable= 9, Autre=98
         electricity_dim = as.factor(ifelse(eclairage %in% c(5, 6, 7, 8, 9, 98), 1, 0))%>% structure(label = "Standard of living: electricity"),
         
  # Creating a dimension for housing quality, checking specific types of construction materials for walls, floor, and roof
        # For nature_mur, 5 is Bois, 6 is "Banco", 7 is "Banco+ enduit ciment",  8 is 'Paille/Tige" and 9 is "Autre".
         # For nature_sol, 3 is 'Argile/banco", 4 is Sable and 7 is "Bois ciré" and 8 is "autre".
         # For nature_toit, 0 is "Autre" and 4 is "Chaume/paille"
         housing_dim = as.factor(ifelse(nature_mur %in% c(5, 6, 7, 8, 9) & nature_sol %in% c(3, 4, 7, 8) & nature_toit %in% c(0, 4), 1, 0))%>% structure(label = "Standard of living: housing"),
         
  # Creating a dimension for household assets, checking if various assets are absent in the household
  
         assets_dim = as.factor(ifelse(radio == 0 & televiseur == 0 & telephone == 0 & telephone_tele == 0 & animaux_traite == 0 & bicyclette == 0 & mobylette == 0 & mobylette_bicycl == 0 & 
                           refri_conge == 0 & refriger_congelat == 0 & voiture == 0 & voiture_cam == 0, 1, 0))%>% structure(label = "Standard of living: assets")
  )



```


```{r}

## In the following we merge the "menage_2002" with the individual dataset
menage_2002 <- menage_2002 %>% 
  left_join(
    
    individus_2002 %>% 
      
      # Filtering to get only individuals marked as "Chef de menage" in QB03
      filter(QB03== 1) %>% #"Chef de menage"
      
      # Renaming the QB15 column to "niveau_instruction_cm"
      mutate(niveau_instruction_cm = QB15 %>% structure(label=Hmisc::label(individus_2002$QB15))) %>% 
  
    rename(
      sexe_cm = QB04,        # Renamed QB04 to "sexe_cm"
      age_cm = QB06,         # Renamed QB06 to "age_cm"
    ) %>% 
      # Selecting specific columns from the filtered dataset
      select(ID_MENAGE, num_men, DR, niveau_instruction_cm, sexe_cm, age_cm, school_attendance_dim, years_schooling_dim),
      
      # Merging by matching columns in both datasets
      by = c("ID_MENAGE" = "ID_MENAGE", 
             "num_men"="num_men",
             "id_dr" = "DR")
    
  )


## In the following we merge the "menage_2013" with the individual dataset
menage_2013 <- menage_2013 %>% 
  left_join(
    
    individus_2013 %>% 
      
      # Filtering to get only individuals marked as "Chef de menage" in B04
      filter(B04==1) %>% #"Chef de ménage"
      
      # Renaming the B32 column to "niveau_instruction_cm"
      mutate(niveau_instruction_cm = B32 %>% structure(label=Hmisc::label(individus_2013$B32))) %>% 
      
       # Rename columns for clarity
  rename(
    sexe_cm = B06,        # Renamed B06 to "sexe_cm"
    age_cm = B08,         # Renamed B08 to "age_cm"
  ) %>% 
      # Selecting specific columns from the filtered dataset
      select(ID_MENAGE, IDDR, niveau_instruction_cm, sexe_cm, age_cm, school_attendance_dim, years_schooling_dim),
      
      # Merging by matching columns in both datasets
      by = c("ID_MENAGE" = "ID_MENAGE", "id_dr" = "IDDR")
    
  )


```


```{r}

# For the RGPH 2013 dataset, recode the "soins_medicaux" variable
# This code modifies the "soins_medicaux" variable, replacing "2" with "0" if present

menage_2013 <- menage_2013 %>%
  mutate(
            soins_medicaux = as.numeric(soins_medicaux),
            soins_medicaux = 
                dplyr::recode(soins_medicaux,"2" = 0),
            soins_medicaux = as.factor(soins_medicaux) %>% structure(label=Hmisc::label(menage_2002$E17))) 

# For the RGPH 2002 dataset, recode the "soins_medicaux" variable
# This code modifies the "soins_medicaux" variable, replacing "2" with "0", "3" with NA, and "9" with NA

menage_2002 <- menage_2002 %>%
  mutate(soins_medicaux = as.numeric(soins_medicaux),
           soins_medicaux = dplyr::recode(soins_medicaux, "2" = 0, "3" = NULL, "9" = NULL), 
         soins_medicaux = as.factor(soins_medicaux) %>% structure(label=Hmisc::label(menage_2002$E17)))

```

```{r wrangling deces}

# For the RGPH 2002 dataset, determine if there was a death in the household in the last twelve months
# This code calculates whether there was a death in the household based on specific columns
deces_2002 <- deces_2002 %>% 
  mutate(
    est_decedee = ifelse(is.na(C03) & is.na(C04) & is.na(C05) & is.na(GROUPE_AGE_DECES), 0, 1) %>% structure(label = "y-a-t-il eu décès dans le ménage au cours des douze derniers mois?")
  ) %>% 
  group_by(ID_MENAGE) %>% 
  summarise(
    
    # Count the number of death in the household
    nbr_deces = sum(est_decedee, na.rm = TRUE),
    
    # Count the number of death in the household below the age of 18
    nbr_deces_under_five = sum(est_decedee==1 & C04 <=18,
                                       na.rm = TRUE)) %>% 
  ungroup() %>% 
  # Create a variable for an under 18 death in the household
  mutate(deces_under_five_dim = as.factor(ifelse(nbr_deces_under_five > 0, 1, ifelse(is.na(nbr_deces_under_five), 0, 0)) %>%
                                            structure(label = "Y a-t-il eu un décés d'un enfant de moins de cinq ans dans le ménage ?")),
         nbr_deces_under_five = nbr_deces_under_five %>%
                                            structure(label = "Nombre de décés de moins de cinq ans")) 

# For the RGPH 2013 dataset, calculate the number of deaths in each household
# This code groups the data by household and calculates the number of deaths in each household
deces_2013 <- deces_2013 %>% 
  group_by(ID_MENAGE) %>% 
  summarise(nbr_deces = sum(C01, na.rm = TRUE)
    
    # Count the number of death in the household below the age of 18
  ) %>% 
  ungroup() %>% 
  # Create the variable for the number of death in the household
  mutate(nbr_deces = as.numeric(nbr_deces) %>%
           structure(label = "Nombre de décés dans le ménage")) 

```


```{r}
# For the RGPH 2013 dataset, select specific columns and merge with "deces_2013" data
# This code selects the "id_dr," "ID_MENAGE," and columns 77 to the last column in "menage_2013"
# It then performs a left join with "deces_2013" based on the "ID_MENAGE" key

# Selecting specific columns from the 'menage_2013' dataset
menage_2013 <- menage_2013 %>% 
  select(id_dr, ID_MENAGE, 77:ncol(menage_2013)) %>%
  
# Performing a left join with the 'deces_2013' dataset based on 'ID_MENAGE'
  left_join(deces_2013, by = "ID_MENAGE") %>%
  
# Adding a new column 'RGPH' with the value "2013" and setting a label
  mutate(RGPH = "2013" %>% structure(label = "RGPH"),
  
# Adding a label to the 'nbr_deces' column
         nbr_deces = nbr_deces %>% structure(label = "Nombre de décès dans le ménage au cours des 12 derniers mois"))

# For the RGPH 2002 dataset, select specific columns and merge with "deces_2013" data
# This code selects the "id_dr," "ID_MENAGE," and columns 71 to the last column in "menage_2002"
# It then performs a left join with "deces_2013" based on the "ID_MENAGE" key

# Selecting specific columns from the 'menage_2002' dataset
menage_2002 <- menage_2002 %>% 
  select(id_dr, ID_MENAGE, num_men, 71:ncol(menage_2002)) %>%
  
# Performing a left join with the 'deces_2002' dataset based on 'ID_MENAGE'
  left_join(deces_2002, by = "ID_MENAGE") %>%
  
# Adding a new column 'RGPH' with the value "2002" and setting a label
  mutate(RGPH = "2002" %>% structure(label = "RGPH"),
  
# Adding a new column 'ID_MENAGE_corrige' by checking if 'num_men' is missing and using 'ID_MENAGE' if so, then setting a label
         ID_MENAGE_corrige = ifelse(is.na(num_men), ID_MENAGE, num_men) %>% structure(label = "ID MENAGE")) %>%
  
# Selecting and removing specific columns
  select(-c(ID_MENAGE, num_men)) %>%
  
# Renaming the 'ID_MENAGE_corrige' column to 'ID_MENAGE'
  rename(ID_MENAGE = ID_MENAGE_corrige) %>%
  
# Adding a new column 'nbr_deces' by checking if it's missing and setting it to 0, with a label
  mutate(nbr_deces = ifelse(is.na(nbr_deces), 0, nbr_deces) %>% structure(label = "Nombre de décès dans le ménage au cours des 12 derniers mois"),
         
         # Replace Na by 0
         
         nbr_deces_under_five = ifelse(is.na(nbr_deces_under_five), 0, nbr_deces_under_five),
         
         deces_under_five_dim = as.factor(ifelse(is.na(deces_under_five_dim), 0, deces_under_five_dim)) %>% structure(label = "Health: child mortality"))

```


```{r}
# Filtering the 'menage_2013' dataset to exclude specific values in the 'id_dr' column
menage_2013 <- menage_2013 %>% 
  filter(!(id_dr %in% c("014301140124", "014301140125")))

# Mutating the 'sp_rgph_2002' dataset to create a new 'id_dr' column based on conditions
sp_rgph_2002 <- sp_rgph_2002 %>% 
  mutate(id_dr = ifelse(NUMDEPT=="3", 
                      paste0(NUM_REG, NUMDEPT, NUM_ARRDT, "18", NUM_CA, IDDR, "O"), 
                      paste0("0", NUM_REG, NUMDEPT, NUM_ARRDT, "8", NUM_CA, IDDR, "O")))

```


```{r}
# Replace 'NA' with 0 (for the nbr deces )

## For RGPH 2013
menage_2013 <- menage_2013 %>%
  
  #replace all NA values with zero
  mutate_if(is.numeric , replace_na, replace = 0)

## For RGPH 2002
menage_2002 <- menage_2002 %>%
  
  #replace all NA values with zero
  mutate_if(is.numeric , replace_na, replace = 0)


# Ranking variables

## For RGPH 2013
menage_2013 <- menage_2013 %>%
 select(id_dr, ID_MENAGE, RGPH, everything())

## For RGPH 2002
menage_2002 <- menage_2002 %>%
 select(id_dr, ID_MENAGE, RGPH, everything())
```



```{r}
# Exporting the "menage_2013" and "menage_2002" datasets

# For the "menage_2002"
# Saving in SPSS format
haven::write_sav(menage_2002,paste0(here::here(), "/output/output_data/menage_2002.sav"))

# Saving in R format
saveRDS(menage_2002,file=paste0(here::here(), "/output/output_data/menage_2002.rds"))

# For the "menage_2013"
# Saving in SPSS format
haven::write_sav(menage_2013,paste0(here::here(), "/output/output_data/menage_2013.sav"))

# Saving in R format
saveRDS(menage_2013,file=paste0(here::here(), "/output/output_data/menage_2013.rds"))


# For the "sp_rgph_2002"
# Saving in the sf format
sf::st_write(obj=sp_rgph_2002, paste0(here::here(), "/output/output_data/EAs_2002_new.shp"),  driver = "ESRI Shapefile", delete_layer = TRUE) # we change the id_dr


```

```{r}
## Removing all objects
rm(list=ls())
```
