---
title: "Wrangling Senegal Census data with tidyverse tools"
author: "IBRAHIM KASSOUM Habibou & HEMA Aboubacar"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
```


## Project Sénégal

### Preambule: 

```{r Package needed, echo = FALSE}

## Importing library
### List of required packages
required_packages <- c("tidyverse","janitor" ,"readr","dplyr","haven","sf",
                       "flextable","sp", "factoextra", "FactoMineR","gtsummary")

# Check if packages are installed
missing_packages <- setdiff(required_packages, installed.packages()[,"Package"])

### Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

### Load all packages
lapply(required_packages, library, character.only = TRUE)

```


### Importing the dataset

```{r}

# Read shapefile data for 2002 and 2013
sp_rgph_2002 <- sf::read_sf(paste0(here::here(),"/data/SHAPEFILE/EAs_2002.shp"))

sp_rgph_2013 <- sf::read_sf(paste0(here::here(),"/data/SHAPEFILE/EAs_2013.shp"))

# Display the first 10 rows of the 2002 and 2013 shapefiles
head(sp_rgph_2002, 10L)
head(sp_rgph_2013, 10L)
```
### Importing the dataset an checking for duplicated row

```{r}
# Read and check for duplicated rows in 2002 datasets

# Read and check for duplicated rows in the "deces_2002" dataset
deces_2002 <- read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 Deces.sav"))
deces_2002 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "individus_2002" dataset
individus_2002 <- haven::read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 Individus.v2.sav"))
individus_2002 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "menage_2002" dataset
menage_2002 <- haven::read_sav(paste0(here::here(),"/data/1. RGPH 2002/Requete PDBA2k2 menage.sav"))
menage_2002 %>% janitor::get_dupes() 
```


```{r}

# Read and check for duplicated rows in 2013 datasets

# Read and check for duplicated rows in the "deces_2013" dataset
deces_2013 <- read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 Deces_Guediawaye.sav"))
deces_2013 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "individus_2013" dataset
individus_2013 <- haven::read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 Indidivus GUEDIAWAYE.sav"))
individus_2013 %>% janitor::get_dupes() 

# Read and check for duplicated rows in the "menage_2013" dataset
menage_2013 <- haven::read_sav(paste0(here::here(),"/data/4. RGPH 2013/Requete PDBA2k13 menage.sav"))
menage_2013 %>% janitor::get_dupes() 
```


```{r}
# Unlabelling the variables in the RGPH datasets
# Transform all "haven_labelled" variables to factor variables

# For 2002
## Unlabel the variables in the "individus_2002" dataset
individus_2002 <- 
  individus_2002 %>% 
  labelled::unlabelled() 

## Unlabel the variables in the "menage_2002" dataset
menage_2002 <- 
  menage_2002 %>% 
  labelled::unlabelled() 

# For 2013  
## Unlabel the variables in the "individus_2013" dataset
individus_2013 <- 
  individus_2013 %>% 
  labelled::unlabelled()  

# #Unlabel the variables in the "menage_2013" dataset
menage_2013 <- 
  menage_2013 %>% 
  labelled::unlabelled() 

# Creating a variable "RGPH" to identify each survey 

# For 2002
## Add a new column "RGPH" with the value "2002" to the "individus_2002" dataset
individus_2002 <- 
  individus_2002 %>% 
  mutate(RGPH = "2002")

# For 2013
## Add a new column "RGPH" with the value "2013" to the "individus_2013" dataset
individus_2013 <- 
  individus_2013 %>% 
  mutate(RGPH = "2013")

```


### -	Calcul des indices sur les variables composants liés au ménage (Age des chefs de ménage, par sexe entre les deux recensements

```{r}
# Select the CM and compute it average age
individus_2013 %>% 
  
  select(B04, B06, B08, RGPH) %>%
  
  ##  Select Household head informations
  filter(B04=="Chef de ménage") %>% # Household head (B04 == 1) 
  
  rename(
    sexe_cm = B06,
    lien_cm = B04,
    age_cm = B08,
    ) %>% 
  
  
  bind_rows(
    
    individus_2002 %>% 
                 
      select(QB03, QB04, QB06, RGPH) %>%
    
    ##  Select Household head informations
    filter(QB03=="Chef de menage") %>% # Household head (B04 == 1) 
    
    rename(
      sexe_cm = QB04,
      lien_cm = QB03,
      age_cm = QB06,
      )
    ) %>% 
  
  # Grouping by sexe (B06)
  group_by(RGPH, sexe_cm) %>% 
  
  # Computing number of HH and average age
  summarise( "Sexe du chef de ménage"= sexe_cm,
            "Nombre d'individus" = n(), 
            "Age moyen" = round(mean(age_cm, na.rm = TRUE), digits = 2)
            ) %>% 
  # Remove the group structure
  ungroup() %>% 
  
  # Selecting the needed columns
  select(-c(sexe_cm)) %>% 
  
  distinct() %>% 
  flextable::regulartable()
```

```{r}
# Select the CM (Chef de ménage) and compute the average age

# Select relevant columns (B04, B06, B08, RGPH) from the "individus_2013" dataset
individus_2013 %>% 
  select(B04, B06, B08, RGPH) %>%

  # Filter to select only records where "B04" (Chef de ménage) equals 1
  # This is done to identify the household head
  filter(B04=="Chef de ménage") %>% # Household head (B04 == 1) 

  # Rename columns for clarity
  rename(
    sexe_cm = B06,        # Renamed B06 to "sexe_cm"
    lien_cm = B04,        # Renamed B04 to "lien_cm"
    age_cm = B08,         # Renamed B08 to "age_cm"
  ) %>% 

  # Combine this filtered data with a similar operation on the "individus_2002" dataset
  bind_rows(
    individus_2002 %>% 
      select(QB03, QB04, QB06, RGPH) %>%
    
    # Filter to select only records where "QB03" (Chef de ménage) equals "Chef de menage"
    filter(QB03=="Chef de menage") %>% # Household head (QB03 == "Chef de menage") 
      
    # Rename columns for consistency
    rename(
      sexe_cm = QB04,        # Renamed QB04 to "sexe_cm"
      lien_cm = QB03,        # Renamed QB03 to "lien_cm"
      age_cm = QB06,         # Renamed QB06 to "age_cm"
    ) %>%
    
    # Replace "Feminin" with "Féminin" in the "sexe_cm" column
    mutate(sexe_cm = ifelse(sexe_cm=="Feminin", "Féminin", "Masculin"))
  ) %>% 

  # Remove the "lien_cm" (link to household head) column
  select(-lien_cm) %>% 
  
  # Generate a summary table using "tbl_summary" for the specified columns
  tbl_summary(
    by = RGPH,  # Group by "RGPH" column
    label = list(sexe_cm ~ "Sexe du chef de ménage",
                 age_cm ~ "Age du chef de ménage (en année)"),
    statistic = list(all_continuous() ~ "{mean}", all_categorical() ~ "{p}"),
    digits = everything() ~ c(0,0),
    missing = "no"
  ) %>% 

  # Modify the table header to provide a descriptive label
  modify_header(label ~ "Caractéristique du Chef de ménage entre les deux recensements") %>%
  
  # Add difference statistics to the table
  add_difference() %>% 
  
  # Add overall statistics to the table
  add_overall() %>%
  
  # Add a count of the observations
  add_n()
```



### -	Calcul multivarié pour des indices de pauvreté (bien des ménages, Niveau d’instruction, accès aux services (bien d’aisance, eau, etc.)


```{r}

# Transform values in the "QB15" column of the "individus_2002" dataset
# The transformation categorizes educational levels into descriptive labels

individus_2002 <- individus_2002 %>% 
  mutate(QB15 = case_when(
    QB15 == 0 ~ "aucune",      # If QB15 is 0, set it to "aucune"
    QB15 == 1 ~ "prescolaire", # If QB15 is 1, set it to "prescolaire"
    QB15 %in% seq(2,7,1) ~ "elementaire", # If QB15 is 2 to 7, set it to "elementaire"
    QB15 %in% seq(8,11,1) ~ "moyen", # If QB15 is 8 to 11, set it to "moyen"
    QB15 %in% seq(12,14,1) ~ "secondaire", # If QB15 is 12 to 14, set it to "secondaire"
    QB15 %in% seq(15,19,1) ~ "superieur", # If QB15 is 15 to 19, set it to "superieur"
    TRUE ~ as.character(QB15) # If none, keep the old value
    ) %>% 
  structure(label = Hmisc::label(individus_2002$QB15))) # Apply labels to the QB15 column


# Transform values in the "B32" column of the "individus_2013" dataset
# The transformation categorizes educational levels from codes into descriptive labels

individus_2013 <- individus_2013 %>% 
  
  mutate(B32 = case_when(
    
    B32 %in% c("Petite section","Moyenne section", "Grnde section") ~ "prescolaire", # If B32 is in pre-school levels, set it to "prescolaire"
    
    B32 %in% c("Cours d'initiation (CI)", "Cours préparatoire (CP)", 
    "Cours élémentaire 1ère année (CE1)", "Cours élémentaire 2ème année (CE2)",
    "Cours moyen 1ère année (CM1)", "Cours moyen 2ème année (CM2)") ~ "elementaire", # If B32 corresponds to elementary levels, set it to "elementaire"
    
    B32 %in% c("Sixième (6ième)", "Cinquième (5ième)", "Quatrième (4ième)", "Troisième (3ième)") ~ "moyen", # If B32 corresponds to middle school levels, set it to "moyen"
    
    B32 %in% c("Seconde", "Première", "Terminale") ~ "secondaire", # If B32 corresponds to high school levels, set it to "secondaire"
    
    B32 %in% c("1ère année", "2ème année", "Licence/3ième année", "Maitrise/4ième année", "DEA/5ième année", "6ième année", "7ième année", "8ième année et +") ~ "superieur", # If B32 corresponds to higher education levels, set it to "superieur"
    TRUE ~ as.character(B32) %>% 
  structure(label = Hmisc::label(individus_2013$B32)))) # Apply labels to the B32 column
```


```{r}

# Renaming variables so that they can have the same name in the two
# datasets

# Renaming variables so that they can have the same name in the two datasets

## For the RGPH 2002
menage_2002 <- menage_2002 %>% 
  
  # Mutate and rename variables in the "menage_2002" dataset
  mutate(
    
    nb_indiv = E01 %>% structure(label=Hmisc::label(menage_2002$Nb_indiv)),
    id_dr = DR %>% structure(label=Hmisc::label(menage_2002$DR)),
    
    type_log = E01 %>% structure(label=Hmisc::label(menage_2002$E01)),
    nbr_piece_occup = E02 %>% structure(label=Hmisc::label(menage_2002$E02)),
    stat_occupation = E03 %>% structure(label=Hmisc::label(menage_2002$E03)),
    type_aisance = E04 %>% structure(label=Hmisc::label(menage_2002$E04)),
    approv_eau = E05 %>% structure(label=Hmisc::label(menage_2002$E05)),
    eclairage = E06 %>% structure(label=Hmisc::label(menage_2002$E06)),
    combustible_cuisine = E07 %>% structure(label=Hmisc::label(menage_2002$E07)),
    
    nature_mur = E08 %>% structure(label=Hmisc::label(menage_2002$E08)),
    nature_toit = E09 %>% structure(label=Hmisc::label(menage_2002$E09)),
    nature_sol = E10 %>% structure(label=Hmisc::label(menage_2002$E10)),
    radio = E11B %>% structure(label=Hmisc::label(menage_2002$E11B)),
    televiseur = E11C %>% structure(label=Hmisc::label(menage_2002$E11C)),
    video = E11D %>% structure(label=Hmisc::label(menage_2002$E11D)),
    refri_conge = E11E %>% structure(label=Hmisc::label(menage_2002$E11E)),
    
    telephone = E11F %>% structure(label=Hmisc::label(menage_2002$E11F)),
    rechaud_cuis = E11G %>% structure(label=Hmisc::label(menage_2002$E11G)),
    foyer_ameli = E11H %>% structure(label=Hmisc::label(menage_2002$E11H)),
    climatisateur = E11I %>% structure(label=Hmisc::label(menage_2002$E11I)),
    machine_coudre = E11J %>% structure(label=Hmisc::label(menage_2002$E11J)),
    aucun_11K = E11K %>% structure(label=Hmisc::label(menage_2002$E11K)),
    voiture = E12B %>% structure(label=Hmisc::label(menage_2002$E12B)),
    
    mobylette = E12C %>% structure(label=Hmisc::label(menage_2002$E12C)),
    bicyclette = E12E %>% structure(label=Hmisc::label(menage_2002$E12E)),
    caleche_charette = E12F %>% structure(label=Hmisc::label(menage_2002$E12F)),
    pirogue = E12G %>% structure(label=Hmisc::label(menage_2002$E12G)),
    aucun_G12 = E12G %>% structure(label=Hmisc::label(menage_2002$E12G)),
    houe_charrue = E13B %>% structure(label=Hmisc::label(menage_2002$E13B)),
    caleche_char = E13C %>% structure(label=Hmisc::label(menage_2002$E13C)),
    
    animaux_traite = E13D %>% structure(label=Hmisc::label(menage_2002$E13D)),
    tracteur = E13E %>% structure(label=Hmisc::label(menage_2002$E13E)),
    voiture_cam = E13F %>% structure(label=Hmisc::label(menage_2002$E13F)),
    mobylette_bicycl = E13G %>%     structure(label=Hmisc::label(menage_2002$E13G)),
    refriger_congelat = E13I %>% structure(label=Hmisc::label(menage_2002$E13I)),
    machine_coudre = E13J %>% structure(label=Hmisc::label(menage_2002$E13J)),
    materiel_musique = E13K %>% structure(label=Hmisc::label(menage_2002$E13K)),
    
    chaise_bache = E13L %>% structure(label=Hmisc::label(menage_2002$E13L)),
    telephone_tele = E13M %>% structure(label=Hmisc::label(menage_2002$E13M)),
    photocopieuse = E13N %>% structure(label=Hmisc::label(menage_2002$E13N)),
    ordinateur = E13O %>% structure(label=Hmisc::label(menage_2002$E13O)),
    moulin = E13P %>% structure(label=Hmisc::label(menage_2002$E13P)),
    appareil_photo = E13Q %>% structure(label=Hmisc::label(menage_2002$E13Q)),
    terrain_bat = E13R %>% structure(label=Hmisc::label(menage_2002$E13R)),
    
    aucun_produc = E13S %>% structure(label=Hmisc::label(menage_2002$E13S)),
    evacut_ordures = E14 %>% structure(label=Hmisc::label(menage_2002$E14)),
    repas_saute = E15 %>% structure(label=Hmisc::label(menage_2002$E15)),
    soins_medicaux = E16 %>% structure(label=Hmisc::label(menage_2002$E16))
  )


# Mutate and rename variables in the "menage_2013" dataset
menage_2013 <- menage_2013 %>% 
  
  mutate(
    id_dr = IDDR %>% structure(label=Hmisc::label(menage_2013$IDDR)),
    nb_indiv = Nbre_indiv %>% structure(label=Hmisc::label(menage_2013$Nbre_indiv)),
    type_log = E01 %>% structure(label=Hmisc::label(menage_2013$E01)),
    nbr_piece_occup = E02 %>% structure(label=Hmisc::label(menage_2013$E02)),
    stat_occupation = E03 %>% structure(label=Hmisc::label(menage_2013$E03)),
    type_aisance = E08 %>% structure(label=Hmisc::label(menage_2013$E08)),
    approv_eau = E09 %>% structure(label=Hmisc::label(menage_2013$E09)),
    eclairage = E11 %>% structure(label=Hmisc::label(menage_2013$E11)),
    combustible_cuisine = E12 %>% structure(label=Hmisc::label(menage_2013$E12)),
    
    nature_mur = E05 %>% structure(label=Hmisc::label(menage_2013$E05)),
    nature_toit = E06 %>% structure(label=Hmisc::label(menage_2013$E06)),
    nature_sol = E07 %>% structure(label=Hmisc::label(menage_2013$E07)),
    radio = E13_1 %>% structure(label=Hmisc::label(menage_2013$E13_1)),
    televiseur = E13_2 %>% structure(label=Hmisc::label(menage_2013$E13_2)),
    video = E13_3 %>% structure(label=Hmisc::label(menage_2013$E13_3)),
    refri_conge = E13_4 %>% structure(label=Hmisc::label(menage_2013$E13_4)),
    
    telephone = E13_6 %>% structure(label=Hmisc::label(menage_2013$E13_6)),
    rechaud_cuis = E13_18 %>% structure(label=Hmisc::label(menage_2013$E13_18)),
    foyer_ameli = E13_7 %>% structure(label=Hmisc::label(menage_2013$E13_7)),
    climatisateur = E13_8 %>% structure(label=Hmisc::label(menage_2013$E13_8)),
    machine_coudre = E13_9 %>% structure(label=Hmisc::label(menage_2013$E13_9)),
    
    voiture = E15_5 %>% structure(label=Hmisc::label(menage_2013$E15_5)),
    mobylette = E15_6 %>% structure(label=Hmisc::label(menage_2013$E15_6)),
    bicyclette = E14_3 %>% structure(label=Hmisc::label(menage_2013$E14_3)),
    caleche_charette = E14_4 %>% structure(label=Hmisc::label(menage_2013$E14_4)),
    pirogue = E15_7 %>% structure(label=Hmisc::label(menage_2013$E15_7)),
    houe_charrue = E15_1 %>% structure(label=Hmisc::label(menage_2013$E15_1)),
    caleche_char = E15_2 %>% structure(label=Hmisc::label(menage_2013$E15_2)),
    
    animaux_traite = E15_3 %>% structure(label=Hmisc::label(menage_2013$E15_3)),
    tracteur = E15_4 %>% structure(label=Hmisc::label(menage_2013$E15_4)),
    voiture_cam = E15_5 %>% structure(label=Hmisc::label(menage_2013$E15_5)),
    mobylette_bicycl = E15_6 %>% structure(label=Hmisc::label(menage_2013$E15_6)),
    refriger_congelat = E15_8 %>% structure(label=Hmisc::label(menage_2013$E15_8)),
    machine_coudre = E15_9 %>% structure(label=Hmisc::label(menage_2013$E15_9)),
    materiel_musique = E15_10 %>% structure(label=Hmisc::label(menage_2013$E15_10)),
    
    chaise_bache = E15_11 %>% structure(label=Hmisc::label(menage_2013$E15_11)),
    telephone_tele = E15_12 %>% structure(label=Hmisc::label(menage_2013$E15_12)),
    photocopieuse = E15_13 %>% structure(label=Hmisc::label(menage_2013$E15_13)),
    ordinateur = E15_14 %>% structure(label=Hmisc::label(menage_2013$E15_14)),
    moulin = E15_15 %>% structure(label=Hmisc::label(menage_2013$E15_15)),
    appareil_photo = E15_16 %>% structure(label=Hmisc::label(menage_2013$E15_16)),
    terrain_bat = E15_17 %>% structure(label=Hmisc::label(menage_2013$E15_17)),
    
    evacut_ordures = E16 %>% structure(label=Hmisc::label(menage_2013$E16)),
    repas_saute = E18 %>% structure(label=Hmisc::label(menage_2013$E18)),
    soins_medicaux = E20 %>% structure(label=Hmisc::label(menage_2013$E20))
    
    )



## In the following we merge the "menage_2002" with the individual dataset
menage_2002 <- menage_2002 %>% 
  left_join(
    
    individus_2002 %>% 
      
      # Filtering to get only individuals marked as "Chef de menage" in QB03
      filter(QB03=="Chef de menage") %>% 
      
      # Renaming the QB15 column to "niveau_instruction_cm"
      rename(niveau_instruction_cm = QB15) %>% 
      
      # Selecting specific columns from the filtered dataset
      select(ID_MENAGE, DR, niveau_instruction_cm),
      
      # Merging by matching columns in both datasets
      by = c("ID_MENAGE" = "ID_MENAGE", "id_dr" = "DR")
    
  )


## In the following we merge the "menage_2013" with the individual dataset
menage_2013 <- menage_2013 %>% 
  left_join(
    
    individus_2013 %>% 
      
      # Filtering to get only individuals marked as "Chef de menage" in B04
      filter(B04=="Chef de ménage") %>% 
      
      # Renaming the B32 column to "niveau_instruction_cm"
      rename(niveau_instruction_cm = B32) %>% 
      
      # Selecting specific columns from the filtered dataset
      select(ID_MENAGE, IDDR, niveau_instruction_cm),
      
      # Merging by matching columns in both datasets
      by = c("ID_MENAGE" = "ID_MENAGE", "id_dr" = "IDDR")
    
  )


```



```{r}
# 
# menage_2002 <- menage_2002 %>% 
#   select(id_dr, ID_MENAGE, 71:118)
# 
# var_tmp<-names(df)
# df <-df %>% dplyr::mutate_at(var_tmp, ~as.numeric(recode(., "Yes"=1, "No"=0, .default = NaN)))
# 
# menage_2013 %>% 
#     mutate_if(across(.cols = everything(), 
#            ~ ifelse(. == "Oui", 1, ifelse(. == "Non"| .=="pas de malade", 0, .))))
# 
# menage_2013 %>% 
#     mutate(across(.cols = where(is.factor), ~ recode(., "Oui"=1, "Non"=0, "pas de malade"= 0, .))) %>% View()
# 
# menage_2013 <- menage_2013 %>% 
#   select(id_dr, ID_MENAGE, 71:118)
```



```{r eval=FALSE, include=FALSE}
 
# the PCA
res.mca <- df_curr %>% 
  select(decbighh_rec, 
         decfemhcare_rec,decfamvisit_rec, 
         dechusearn_rec, wkearnsmore_rec, decfemearn_rec) %>% 
  FactoMineR::MCA(ncp = 5, graph = TRUE, method="Burt", na.method = "Average")

# Eigen values

eig.val <- get_eigenvalue(res.mca)
fviz_screeplot(res.mca, addlabels = TRUE, ylim = c(0, 80))

# Correlation between variables and Principal dimensions
fviz_mca_var(res.mca, col.var = "contrib", 
             repel = TRUE, # Avoid text overlapping (slow)
             choice = c("var.cat"),
             ggtheme = theme_minimal()) +
  scale_color_gradient2(low = "white", mid = "grey",
                        high = "black", midpoint = 2)
ggsave("reg_tables/TotalEvent/conflict_post/default/variable_graphique.jpeg")

# fviz_mca_ind(res.mca, 
#              label = "none", # hide individual labels
#              habillage = "decbighh_rec", # color by groups 
#             # palette = c("#00AFBB", "#E7B800","#8fd175"),
#              addEllipses = TRUE, ellipse.type = "confidence",
#              ggtheme = theme_minimal()) 

# fviz_mca_biplot(res.mca, 
#                  repel = TRUE, # Avoid text overlapping (slow if many point)
#                  ggtheme = theme_minimal())

var <- get_mca_var(res.mca)
var
round(var$coord, 2)



# Get the individual coordinates (loadings) for each observation
individual_coordinates <- res.mca$ind$coord

# affect the first dimension of the dataframe
df_curr$burt_mat <- individual_coordinates[,1]
```


### -	Calcul indice sur la dynamique de la population (croissance démographique au niveau des DR)



