---
title: "Spatial analysis using the Guediawaye Census data"
author: "HEMA Aboubacar"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
```



```{r}
rm(list=ls())
```

## Importing library

```{r Package needed, results = "hide"}


## Importing library
### List of required packages
required_packages <- c("tidyverse","janitor" ,"readr","dplyr","haven","sf", "flextable","sp", "factoextra", "FactoMineR","gtsummary", "sjPlot", "fastDummies","ggthemes","spdep")

# Check if packages are installed
missing_packages <- setdiff(required_packages, installed.packages()[,"Package"])

### Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

### Load all packages
lapply(required_packages, library, character.only = TRUE)

```



```{r}
# Read shapefile data for 2002 and 2013


MPI_data_dr_2002 <- sf::read_sf(paste0(here::here(),"/output/output_data/MPI_data_dr_2002.shp"))
MPI_data_dr_2013 <- sf::read_sf(paste0(here::here(),"/output/output_data/MPI_data_dr_2013.shp"))
```


```{r}
neighbors_2013 <- poly2nb(MPI_data_dr_2013, queen = TRUE)
neighbors_2002 <- poly2nb(MPI_data_dr_2002, queen = TRUE)

```

```{r}
dfw_coords <- MPI_data_dr_2013 %>%
  st_centroid() %>%
  st_coordinates()

plot(MPI_data_dr_2013$geometry)
plot(neighbors_2013, 
     coords = dfw_coords, 
     add = TRUE, 
     col = "blue", 
     points = FALSE,
     main = "Visualization of queens-case neighborhood relationships")

# ggsave(paste0(here::here(),"/output/output_img/queens-case_neighborhood_relationships_rgph_2013.pdf"), width = 8, height = 5)

dfw_coords <- MPI_data_dr_2002 %>%
  st_centroid() %>%
  st_coordinates()

plot(MPI_data_dr_2002$geometry)
plot(neighbors_2002, 
     coords = dfw_coords, 
     add = TRUE, 
     col = "blue", 
     points = FALSE,
     main = "Visualization of queens-case neighborhood relationships")
```

```{r}
#Generating the spatial weights matrix
weights_2013 <- nb2listw(neighbors_2013, style = "W")
weights_2002 <- nb2listw(neighbors_2002, style = "W")
```

# Global and local spatial autocorrelation

## Spatial lags and Moran’s I



```{r}
MPI_data_dr_2013$lag_MPI_men <- lag.listw(weights_2013, MPI_data_dr_2013$MPI_men)
MPI_data_dr_2002$lag_MPI_men <- lag.listw(weights_2002, MPI_data_dr_2002$MPI_men)
```

```{r}
ggplot(MPI_data_dr_2013, aes(x = MPI_men, y = lag_MPI_men)) + 
  geom_point(alpha = 0.3) + 
  geom_abline(color = "red") + 
  theme_minimal() + 
  labs(title = " Scatterplot of MPI relative to its spatial lag for Guediawaye census 2013",
       x = "MPI",
       y = "Spatial lag, MPI", 
       caption = "")
ggsave(paste0(here::here(),"/output/output_img/mpi_spatialLag_rgph_2013.pdf"), width = 8, height = 5)

ggplot(MPI_data_dr_2002, aes(x = MPI_men, y = lag_MPI_men)) + 
  geom_point(alpha = 0.3) + 
  geom_abline(color = "red") + 
  theme_minimal() + 
  labs(title = " Scatterplot of MPI relative to its spatial lag for Guediawaye census 2002",
       x = "MPI",
       y = "Spatial lag, MPI", 
       caption = "")
ggsave(paste0(here::here(),"/output/output_img/mpi_spatialLag_rgph_2002.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/mpi_spatialLag_rgph_2002.png"), width = 8, height = 5)
```
```{r}
moran.test(MPI_data_dr_2013$MPI_men, weights_2013)
moran.test(MPI_data_dr_2002$MPI_men, weights_2002)
```


## Local spatial autocorrelation

We can explore this further with local spatial autocorrelation analysis. Local measures of spatial autocorrelation disaggregate global results to identify “hot spots” of similar values within a given spatial dataset. One such example is the Getis-Ord local G statistic (Getis and Ord 1992), which is computed as follows:


$$
G_i = \frac{\sum_j{w_{ij}x_j}}{\sum_{j=1}^nx_j} 
$$




Given that the returned results are z-scores, an analyst can choose hot spot thresholds in the statistic, calculate them with case_when(), then plot them accordingly.



```{r echo=TRUE}
# For Gi*, re-compute the weights with `include.self()`
localg_weights_2013 <- nb2listw(include.self(neighbors_2013))
MPI_data_dr_2013$localG <- localG(MPI_data_dr_2013$MPI_men, localg_weights_2013)
MPI_data_dr_2013$localG <- as.numeric(MPI_data_dr_2013$localG)
#
localg_weights_2002 <- nb2listw(include.self(neighbors_2002))
MPI_data_dr_2002$localG <- localG(MPI_data_dr_2002$MPI_men, localg_weights_2002)
MPI_data_dr_2002$localG <- as.numeric(MPI_data_dr_2002$localG)
```

```{r}
ggplot(MPI_data_dr_2013) + 
  geom_sf(aes(fill = localG), color = NA) + 
  scale_fill_distiller(palette = "RdYlBu") + 
  theme_void() + 
  labs(fill = "Local Gi* statistic") +
labs(title = "Map of local Gi* scores for Guediawaye Census 2013")
ggsave(paste0(here::here(),"/output/output_img/localgi_rgph_2013.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgi_rgph_2013.png"), width = 8, height = 5)
#
ggplot(MPI_data_dr_2002) + 
  geom_sf(aes(fill = localG), color = NA) + 
  scale_fill_distiller(palette = "RdYlBu") + 
  theme_void() + 
  labs(fill = "Local Gi* statistic") +
labs(title = "Map of local Gi* scores for Guediawaye Census 2002")
ggsave(paste0(here::here(),"/output/output_img/localgi_rgph_2002.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgi_rgph_2002.png"), width = 8, height = 5)
```


```{r}
df <- MPI_data_dr_2002 %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(MPI_data_dr_2013 %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


df %>% 
    
      ggplot(aes(fill = localG)) + 
      
  geom_sf(size = 0.1) + 
  scale_fill_distiller(palette = "RdYlBu") +
  #scale_fill_brewer(palette = "Set2") + 
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            #legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            #panel.grid = element_line(color = "white", size = 0.8),
            legend.box.background = element_rect(),
            legend.box.margin = margin(6, 6, 6, 6),
            plot.background = element_rect(fill = "white"))+
  ggspatial::annotation_scale(
    location = "br",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tl", which_north = "true",
    #pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
    #which_north = "grid",
  height = unit(1, "cm"),
  width = unit(1, "cm"),
  )

#ggsave(paste0(here::here(),"/output/output_img/lisacluster.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgi.png"), width = 8, height = 5)
```






Given that the returned results are z-scores, an analyst can choose hot spot thresholds in the statistic, calculate them with case_when(), then plot them accordingly.

```{r}
MPI_data_dr_2013 <- MPI_data_dr_2013 %>%
  mutate(hotspot = case_when(
    localG >= 2.576 ~ "High cluster",
    localG <= -2.576 ~ "Low cluster",
    TRUE ~ "Not significant"
  ))

ggplot(MPI_data_dr_2013) + 
  geom_sf(aes(fill = hotspot), color = "grey90", size = 0.1) + 
  scale_fill_manual(values = c("red", "blue", "grey")) + 
  theme_void() +
labs(title = "Map of local Gi* scores with significant clusters highlighted for Guediawaye Census 2013")
ggsave(paste0(here::here(),"/output/output_img/localgicluster_rgph_2013.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgicluster_rgph_2013.png"), width = 8, height = 5)
###
MPI_data_dr_2002 <- MPI_data_dr_2002 %>%
  mutate(hotspot = case_when(
    localG >= 2.576 ~ "High cluster",
    localG <= -2.576 ~ "Low cluster",
    TRUE ~ "Not significant"
  ))

ggplot(MPI_data_dr_2002) + 
  geom_sf(aes(fill = hotspot), color = "grey90", size = 0.1) + 
  scale_fill_manual(values = c("red", "blue", "grey")) + 
  theme_void() +
labs(title = "Map of local Gi* scores with significant clusters highlighted for Guediawaye Census 2002")
ggsave(paste0(here::here(),"/output/output_img/localgicluster_rgph_2002.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgicluster_rgph_2002.png"), width = 8, height = 5)
```
```{r}
df <- MPI_data_dr_2002 %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(MPI_data_dr_2013 %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


df %>% 
    
      ggplot(aes(fill = hotspot)) + 
      
  geom_sf(size = 0.1) + 
  scale_fill_manual(values = c("red", "blue", "grey")) +
  #scale_fill_brewer(palette = "Set2") + 
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            #legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            #panel.grid = element_line(color = "white", size = 0.8),
            legend.box.background = element_rect(),
            legend.box.margin = margin(6, 6, 6, 6),
            plot.background = element_rect(fill = "white"))+
  ggspatial::annotation_scale(
    location = "br",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tl", which_north = "true",
    #pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
    #which_north = "grid",
  height = unit(1, "cm"),
  width = unit(1, "cm"),
  )

#ggsave(paste0(here::here(),"/output/output_img/lisacluster.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/localgicluster.png"), width = 8, height = 5)
```



## Identifying clusters and spatial outliers with local indicators of spatial association (LISA)

```{r}
set.seed(1994)

MPI_data_dr_2013$scaled_MPI_men <- as.numeric(scale(MPI_data_dr_2013$MPI_men))
MPI_data_dr_2002$scaled_MPI_men <- as.numeric(scale(MPI_data_dr_2002$MPI_men))

dfw_lisa_2013 <- localmoran_perm(
  MPI_data_dr_2013$scaled_MPI_men, 
  weights_2013, 
  nsim = 999L, 
  alternative = "two.sided"
) %>%
  as_tibble() %>%
  set_names(c("local_i", "exp_i", "var_i", "z_i", "p_i",
              "p_i_sim", "pi_sim_folded", "skewness", "kurtosis"))

dfw_lisa_2013_df <- MPI_data_dr_2013 %>%
  select(id_dr, scaled_MPI_men) %>%
  mutate(lagged_estimate = lag.listw(weights_2013, scaled_MPI_men)) %>%
  bind_cols(dfw_lisa_2013)

##
dfw_lisa_2002 <- localmoran_perm(
  MPI_data_dr_2002$scaled_MPI_men, 
  weights_2002, 
  nsim = 999L, 
  alternative = "two.sided"
) %>%
  as_tibble() %>%
  set_names(c("local_i", "exp_i", "var_i", "z_i", "p_i",
              "p_i_sim", "pi_sim_folded", "skewness", "kurtosis"))

dfw_lisa_2002_df <- MPI_data_dr_2002 %>%
  select(id_dr, scaled_MPI_men) %>%
  mutate(lagged_estimate = lag.listw(weights_2002, scaled_MPI_men)) %>%
  bind_cols(dfw_lisa_2002)
```

The above code uses the following steps:

First, a random number seed is set given that we are using the conditional permutation approach to calculating statistical significance. This will ensure reproducibility of results when the process is re-run.
The ACS estimate for median age is converted to a z-score using scale(), which subtracts the mean from the estimate then divides by its standard deviation. This follows convention from GeoDa.
LISA is computed with localmoran_perm() for the scaled value for median age, using the contiguity-based spatial weights matrix. 999 conditional permutation simulations are used to calculate statistical significance, and the argument alternative = "two.sided" will identify both statistically significant clusters and statistically significant spatial outliers.
The LISA data frame is attached to the Census tract shapes after computing the lagged value for median age.


```{r}
dfw_lisa_2013_clusters <- dfw_lisa_2013_df %>%
  mutate(lisa_cluster = case_when(
    p_i >= 0.05 ~ "Not significant",
    scaled_MPI_men > 0 & local_i > 0 ~ "High-high",
    scaled_MPI_men > 0 & local_i < 0 ~ "High-low",
    scaled_MPI_men < 0 & local_i > 0 ~ "Low-low",
    scaled_MPI_men < 0 & local_i < 0 ~ "Low-high"
  ))


##
dfw_lisa_2002_clusters <- dfw_lisa_2002_df %>%
  mutate(lisa_cluster = case_when(
    p_i >= 0.05 ~ "Not significant",
    scaled_MPI_men > 0 & local_i > 0 ~ "High-high",
    scaled_MPI_men > 0 & local_i < 0 ~ "High-low",
    scaled_MPI_men < 0 & local_i > 0 ~ "Low-low",
    scaled_MPI_men < 0 & local_i < 0 ~ "Low-high"
  ))
```

The LISA quadrant plot then appears as follow:

```{r}
color_values <- c(`High-high` = "red", 
                  `High-low` = "pink", 
                  `Low-low` = "blue", 
                  `Low-high` = "lightblue", 
                  `Not significant` = "white")

ggplot(dfw_lisa_2013_clusters, aes(x = scaled_MPI_men, 
                              y = lagged_estimate,
                              fill = lisa_cluster)) + 
  geom_point(color = "black", shape = 21, size = 2) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "dashed") + 
  scale_fill_manual(values = color_values) + 
  labs(x = "MPI (z-score)",
       y = "Spatial lag of MPI (z-score)",
       fill = "Cluster type") + 
  labs(title = "LISA quadrant scatterplot for Guediawaye Census 2013")
ggsave(paste0(here::here(),"/output/output_img/lisaquadrant_rgph_2013.pdf"), width = 8, height = 5)

ggsave(paste0(here::here(),"/output/output_img/lisaquadrant_rgph_2013.png"), width = 8, height = 5)

##
ggplot(dfw_lisa_2002_clusters, aes(x = scaled_MPI_men, 
                                   y = lagged_estimate,
                                   fill = lisa_cluster)) + 
  geom_point(color = "black", shape = 21, size = 2) + 
  theme_minimal() + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  geom_vline(xintercept = 0, linetype = "dashed") + 
  scale_fill_manual(values = color_values) + 
  labs(x = "MPI (z-score)",
       y = "Spatial lag of MPI (z-score)",
       fill = "Cluster type") + 
  labs(title = "LISA quadrant scatterplot for Guediawaye Census 2002")
ggsave(paste0(here::here(),"/output/output_img/lisaquadrant_rgph_2002.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/lisaquadrant_rgph_2002.png"), width = 8, height = 5)
```

```{r}

ggplot(dfw_lisa_2013_clusters, aes(fill = lisa_cluster)) + 
  labs(title = " LISA cluster map for Guediawaye Census 2013") +
  geom_sf(size = 0.1) + 
  theme_void() + 
  scale_fill_manual(values = color_values) + 
  labs(fill = "Cluster type")
ggsave(paste0(here::here(),"/output/output_img/lisacluster_rgph_2013.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/lisacluster_rgph_2013.png"), width = 8, height = 5)

###
ggplot(dfw_lisa_2002_clusters, aes(fill = lisa_cluster)) + 
  labs(title = " LISA cluster map for Guediawaye Census 2002") +
  geom_sf(size = 0.1) + 
  theme_void() + 
  scale_fill_manual(values = color_values) + 
  labs(fill = "Cluster type")
ggsave(paste0(here::here(),"/output/output_img/lisacluster_rgph_2002.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/lisacluster_rgph_2002.png"), width = 8, height = 5)
```

```{r}
dfw_lisa_clusters <- dfw_lisa_2002_clusters %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(dfw_lisa_2013_clusters %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


dfw_lisa_clusters %>% 
    
      ggplot(aes(fill = lisa_cluster)) + 
      
  geom_sf(size = 0.1) + 
  scale_fill_manual(values = color_values) +
  #scale_fill_brewer(palette = "Set2") + 
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            #legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            #panel.grid = element_line(color = "white", size = 0.8),
            legend.box.background = element_rect(),
            legend.box.margin = margin(6, 6, 6, 6),
            plot.background = element_rect(fill = "white"))+
  ggspatial::annotation_scale(
    location = "br",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tl", which_north = "true",
    #pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
    #which_north = "grid",
  height = unit(1, "cm"),
  width = unit(1, "cm"),
  )

#ggsave(paste0(here::here(),"/output/output_img/lisacluster.pdf"), width = 8, height = 5)
ggsave(paste0(here::here(),"/output/output_img/lisacluster.png"), width = 8, height = 5)
```

