---
title: "Modeling Guediawaye Census data "
author: "HEMA Aboubacar"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
```



```{r}
rm(list=ls())
```

## Importing library

```{r Package needed, results = "hide"}


## Importing library
### List of required packages
required_packages <- c("tidyverse","janitor" ,"readr","dplyr","haven","sf", "flextable","sp", "factoextra", "FactoMineR","gtsummary", "sjPlot", "fastDummies","ggthemes","spdep","patchwork")

# Check if packages are installed
missing_packages <- setdiff(required_packages, installed.packages()[,"Package"])

### Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

### Load all packages
lapply(required_packages, library, character.only = TRUE)

```


```{r}
# Read shapefile data for 2013 and 2013


MPI_data_dr_2013 <- sf::read_sf(paste0(here::here(),"/output/output_data/MPI_data_dr_2013.shp"))
MPI_data_dr_2002 <- sf::read_sf(paste0(here::here(),"/output/output_data/MPI_data_dr_2002.shp"))
MPI_data_dr <- MPI_data_dr_2013 %>%
  plyr::rbind.fill(MPI_data_dr_2002) %>% 
  st_as_sf()
```


# Regression modeling with Guediawaye Census data

```{r}
#"hh_size",
predictors = c("menage","Aucun","Primair","Moyen","Secondr","fertlty","Ltrcy_F","nbr_cm_h","nbr_cm_f","pct_cm_f")
outcome = "MPI_men"
```

## Inspecting the outcome variable (MPI) with visualization

```{r}

mhv_map <- ggplot(MPI_data_dr_2013, aes(fill = MPI_men)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "MPI ")

mhv_histogram <- ggplot(MPI_data_dr_2013, aes(x = MPI_men)) + 
  geom_histogram(alpha = 0.5, fill = "navy", color = "navy",
                 bins = 100) + 
  theme_minimal() + 
  scale_x_continuous(labels = scales::label_number_si(accuracy = 0.1)) + 
  labs(x = "MPI")

mhv_map + mhv_histogram + labs(title = "MPI value charts for Guediawaye Census 2013")
```


```{r}
MPI_data_dr %>% 
    
      ggplot(aes(fill = MPI_men)) + 
      
      geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "MPI ")+
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```
```{r}
ggplot(MPI_data_dr, aes(x = MPI_men)) + 
  geom_histogram(alpha = 0.5, fill = "navy", color = "navy",
                 bins = 100) + 
  theme_minimal() + 
  scale_x_continuous(labels = scales::label_number_si(accuracy = 0.1)) + 
  labs(x = "MPI")+
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```


```{r}

mhv_map_log <- ggplot(MPI_data_dr_2013, aes(fill = log(MPI_men))) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "MPI\nvalue (log)")

mhv_histogram_log <- ggplot(MPI_data_dr_2013, aes(x = log(MPI_men))) + 
  geom_histogram(alpha = 0.5, fill = "navy", color = "navy",
                 bins = 100) + 
  theme_minimal() + 
  scale_x_continuous() + 
  labs(x = "MPI (log)")

mhv_map_log + mhv_histogram_log + labs(title = "Logged MPI value charts for Guediawaye Census 2013")
```


```{r}
MPI_data_dr %>% 
    
      ggplot(aes(fill = log(MPI_men))) + 
      
      geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "MPI\nvalue (log)")+
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```


## A first regression model

```{r}
library(sf)
library(units)
predictors = c("menage","Primair","Moyen","Secondr","Ltrcy_F","pct_cm_f","hh_size","Ltrcy_A","Ltrcy_W","Ltrcy_M","fertlty")
outcome = "MPI_men"
MPI_data_dr_2013_for_model<- MPI_data_dr_2013 %>%
  dplyr::select(MPI_men,predictors) %>% 
  mutate(pop_density = as.numeric(set_units(hh_size / st_area(.), "1/km2"))) %>% 
  dplyr::select(-hh_size)

##
MPI_data_dr_2002_for_model<- MPI_data_dr_2002 %>%
  dplyr::select(MPI_men,predictors) %>% 
  mutate(pop_density = as.numeric(set_units(hh_size / st_area(.), "1/km2"))) %>% 
  dplyr::select(-hh_size)
```

```{r}

formula <- "log(MPI_men) ~ menage  + Primair + Moyen + Secondr + Ltrcy_F + pct_cm_f + pop_density + Ltrcy_A + Ltrcy_W + Ltrcy_M + fertlty"

model_2013 <- lm(formula = formula, data = MPI_data_dr_2013_for_model)
summary(model_2013)
###
model_2002 <- lm(formula = formula, data = MPI_data_dr_2002_for_model)

summary(model_2002)
```

```{r}
library(corrr)

dfw_estimates_2013 <- MPI_data_dr_2013_for_model%>%
  select(-MPI_men) %>%
  st_drop_geometry()

correlations <- correlate(dfw_estimates_2013, method = "pearson")
network_plot(correlations) + labs(title = "Network plot of correlations between model predictors for Guediawaye Census 2013")

##
dfw_estimates_2002 <- MPI_data_dr_2002_for_model%>%
  select(-MPI_men) %>%
  st_drop_geometry()

correlations <- correlate(dfw_estimates_2002, method = "pearson")
network_plot(correlations) + labs(title = "Network plot of correlations between model predictors for Guediawaye Census 2002")
```

```{r}
library(car)

vif(model_2013)
vif(model_2002)
```

## Dimension reduction with principal components analysis


```{r}
pca_2013 <- prcomp(
  formula = ~., 
  data = dfw_estimates_2013, 
  scale. = TRUE, 
  center = TRUE
)

summary(pca_2013)
##
pca_2002 <- prcomp(
  formula = ~., 
  data = dfw_estimates_2002, 
  scale. = TRUE, 
  center = TRUE
)

summary(pca_2002)
```

```{r}
pca_2013_tibble <- pca_2013$rotation %>%
  as_tibble(rownames = "predictor")
pca_2013_tibble

###
pca_2002_tibble <- pca_2002$rotation %>%
  as_tibble(rownames = "predictor")
pca_2002_tibble
```

```{r}

pca_2013_tibble %>%
  select(predictor:PC5) %>%
  pivot_longer(PC1:PC5, names_to = "component", values_to = "value") %>%
  ggplot(aes(x = value, y = predictor)) + 
  geom_col(fill = "darkgreen", color = "darkgreen", alpha = 0.5) + 
  facet_wrap(~component, nrow = 1) + 
  labs(y = NULL, x = "Value", title = " Loadings for first five principal components  for Guediawaye Census 2013") + 
  theme_minimal()
```

```{r}
pca_2002_tibble %>%
  select(predictor:PC5) %>%
  pivot_longer(PC1:PC5, names_to = "component", values_to = "value") %>%
  ggplot(aes(x = value, y = predictor)) + 
  geom_col(fill = "darkgreen", color = "darkgreen", alpha = 0.5) + 
  facet_wrap(~component, nrow = 1) + 
  labs(y = NULL, x = "Value", title = " Loadings for first five principal components  for Guediawaye Census 2002") + 
  theme_minimal()
```


```{r eval=FALSE, include=FALSE}
pca_2013_tibble<-pca_2013_tibble %>% 
  select(predictor:PC5) %>%
  pivot_longer(PC1:PC5, names_to = "component", values_to = "value")


pca_2013_tibble$RGPH<-"Census 2013"
pca_2002_tibble<-pca_2002_tibble %>% 
  select(predictor:PC5) %>%
  pivot_longer(PC1:PC5, names_to = "component", values_to = "value")
pca_2002_tibble$RGPH<-"Census 2002"

pca_tibble <- pca_2013_tibble %>%
  plyr::rbind.fill(pca_2002_tibble)
esquisse::esquisser(pca_tibble)
pca_tibble %>% 
  ggplot(aes(x = value, y = predictor)) + 
  geom_col(fill = "darkgreen", color = "darkgreen", alpha = 0.5) + 
  facet_wrap(~component, nrow = 1) +
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```

```{r eval=FALSE, include=FALSE}
ggplot(pca_tibble) +
  aes(x = predictor, colour = RGPH, weight = value) +
  geom_bar(position = "dodge", fill = "#112446") +
  scale_color_brewer(palette = "Set1", direction = 1) +
  coord_flip() +
  ggthemes::theme_igray() +
  facet_wrap(vars(component), scales = "free")
```


```{r}
components_2013 <- predict(pca_2013, dfw_estimates_2013)

dfw_pca_2013 <- MPI_data_dr_2013_for_model%>%
  select(MPI_men) %>%
  cbind(components_2013) 

ggplot(dfw_pca_2013, aes(fill = PC1)) +
  geom_sf(color = NA) +
  labs(title = "Map of principal component 1 for Guediawaye Census 2013") +
  theme_void() +
  scale_fill_viridis_c()
```

```{r}
components_2002 <- predict(pca_2002, dfw_estimates_2002)

dfw_pca_2002 <- MPI_data_dr_2002_for_model%>%
  select(MPI_men) %>%
  cbind(components_2002) 

ggplot(dfw_pca_2002, aes(fill = PC1)) +
  geom_sf(color = NA) +
  labs(title = "Map of principal component 1 for Guediawaye Census 2002") +
  theme_void() +
  scale_fill_viridis_c()
```
```{r}
dfw_pca_2002$RGPH<-"Census 2002"
dfw_pca_2013$RGPH<-"Census 2013"
dfw_pca <- dfw_pca_2002 %>%
  plyr::rbind.fill(dfw_pca_2013) %>% 
  st_as_sf()


dfw_pca %>% 
    
      ggplot(aes(fill = PC1))  + 
      
      # Adding spatial features to the plot
      geom_sf(color = NA) +
       scale_fill_viridis_c()+
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```

```{r}
dfw_pca %>% 
    
      ggplot(aes(fill = PC2))  + 
      
      # Adding spatial features to the plot
      geom_sf(color = NA) +
       scale_fill_viridis_c()+
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```




```{r}
pca_2013_formula <- paste0("log(MPI_men) ~ ", 
                      paste0('PC', 1:6, collapse = ' + '))

pca_2013_model <- lm(formula = pca_2013_formula, data = dfw_pca_2013)

summary(pca_2013_model)

###
pca_2002_formula <- paste0("log(MPI_men) ~ ", 
                      paste0('PC', 1:6, collapse = ' + '))

pca_2002_model <- lm(formula = pca_2002_formula, data = dfw_pca_2002)

summary(pca_2002_model)
```

## Spatial regression

```{r}
MPI_data_dr_2013_for_model$residuals <- residuals(model_2013)

ggplot(MPI_data_dr_2013_for_model, aes(x = residuals)) + 
  geom_histogram(bins = 100, alpha = 0.5, color = "navy",
                 fill = "navy") + 
  labs(title = "Distribution of model residuals for Guediawaye Census 2013") +
  theme_minimal()
```

```{r}
MPI_data_dr_2002_for_model$residuals <- residuals(model_2002)

ggplot(MPI_data_dr_2002_for_model, aes(x = residuals)) + 
  geom_histogram(bins = 100, alpha = 0.5, color = "navy",
                 fill = "navy") + 
  labs(title = "Distribution of model residuals for Guediawaye Census 2002") +
  theme_minimal()
```

```{r}
MPI_data_dr_for_model <- MPI_data_dr_2002_for_model %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(MPI_data_dr_2013_for_model %>%
  dplyr::mutate(RGPH = "Census 2013"))

ggplot(MPI_data_dr_for_model, aes(x = residuals)) + 
      geom_histogram(bins = 100, alpha = 0.5, color = "navy",
                 fill = "navy") +
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```


```{r}
library(spdep)

wts_2013 <- MPI_data_dr_2013_for_model %>%
  poly2nb() %>%
  nb2listw()

moran.test(MPI_data_dr_2013_for_model$residuals, wts_2013)
```

```{r}
wts_2002 <- MPI_data_dr_2002_for_model %>%
  poly2nb() %>%
  nb2listw()

moran.test(MPI_data_dr_2002_for_model$residuals, wts_2002)
```


```{r}
MPI_data_dr_2013_for_model$lagged_residuals <- lag.listw(wts_2013, MPI_data_dr_2013_for_model$residuals)

ggplot(MPI_data_dr_2013_for_model, aes(x = residuals, y = lagged_residuals)) + 
  theme_minimal() + 
  labs(title = "Moran scatterplot of residual spatial autocorrelation for Guediawaye Census 2013") +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red")
```
```{r}
MPI_data_dr_2002_for_model$lagged_residuals <- lag.listw(wts_2002, MPI_data_dr_2002_for_model$residuals)

ggplot(MPI_data_dr_2002_for_model, aes(x = residuals, y = lagged_residuals)) + 
  theme_minimal() + 
  labs(title = "Moran scatterplot of residual spatial autocorrelation for Guediawaye Census 2002") +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red")
```
```{r}
MPI_data_dr_for_model <- MPI_data_dr_2002_for_model %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(MPI_data_dr_2013_for_model %>%
  dplyr::mutate(RGPH = "Census 2013"))

ggplot(MPI_data_dr_for_model, aes(x = residuals, y = lagged_residuals)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", color = "red")+
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```



# Geographically weighted regression

## Choosing a bandwidth for GWR

```{r}
library(GWmodel)
library(sf)

dfw_data_sp_2013 <- MPI_data_dr_2013_for_model%>%
  as_Spatial()

bw <- bw.gwr(
  formula = formula, 
  data = dfw_data_sp_2013, 
  kernel = "bisquare",
  adaptive = TRUE
)

###
dfw_data_sp_2002 <- MPI_data_dr_2002_for_model%>%
  as_Spatial()

bw <- bw.gwr(
  formula = formula, 
  data = dfw_data_sp_2002, 
  kernel = "bisquare",
  adaptive = TRUE
)
```

```{r}

gw_model_2013 <- gwr.basic(
  formula = formula, 
  data = dfw_data_sp_2013, 
  bw = bw,
  kernel = "bisquare",
  adaptive = TRUE
)

###
gw_model_2002 <- gwr.basic(
  formula = formula, 
  data = dfw_data_sp_2002, 
  bw = bw,
  kernel = "bisquare",
  adaptive = TRUE
)
```

```{r}
names(gw_model_2013)

##
names(gw_model_2002)
```

```{r}
gw_model_2013_results <- gw_model_2013$SDF %>%
  st_as_sf() 

names(gw_model_2013_results)

###
gw_model_2002_results <- gw_model_2002$SDF %>%
  st_as_sf() 

names(gw_model_2002_results)
```

```{r}
ggplot(gw_model_2013_results, aes(fill = Local_R2)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local R-squared values from the GWR model for Guediawaye Census 2013") +
  theme_void()
```
```{r}
ggplot(gw_model_2002_results, aes(fill = Local_R2)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local R-squared values from the GWR model for Guediawaye Census 2002") +
  theme_void()
```

```{r}
gw_model_results <- gw_model_2002_results %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(gw_model_2013_results %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


gw_model_results %>% 
    
      ggplot(aes(fill = Local_R2)) + 
      
      # Adding spatial features to the plot
      geom_sf() +
      
      scale_fill_viridis_c() +
       
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```





```{r}
ggplot(gw_model_2013_results, aes(fill = menage)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local parameter estimates for household members for Guediawaye Census 2013") +
  theme_void() + 
  labs(fill = "Local β for \nHH")
```
```{r}
ggplot(gw_model_2002_results, aes(fill = menage)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local parameter estimates for household members for Guediawaye Census 2002") +
  theme_void() + 
  labs(fill = "Local β for \nHH")
```

```{r}
gw_model_results <- gw_model_2002_results %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(gw_model_2013_results %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


gw_model_results %>% 
    
      ggplot(aes(fill = menage)) + 
      
      # Adding spatial features to the plot
      geom_sf() +
      
      scale_fill_viridis_c() +
  labs(fill = "Local β for \nHH")+
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```



```{r}
ggplot(gw_model_2013_results, aes(fill = pop_density)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local parameter estimates for population density for Guediawaye Census 2013") +
  theme_void() + 
  labs(fill = "Local β for \npopulation density")
```
```{r}
ggplot(gw_model_2002_results, aes(fill = pop_density)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  labs(title = "Local parameter estimates for population density for Guediawaye Census 2002") +
  theme_void() + 
  labs(fill = "Local β for \npopulation density")
```

```{r}
gw_model_results <- gw_model_2002_results %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(gw_model_2013_results %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


gw_model_results %>% 
    
      ggplot(aes(fill = pop_density)) + 
      
      # Adding spatial features to the plot
      geom_sf() +
      
      scale_fill_viridis_c() +
  labs(fill = "Local β for \npopulation density")+
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```



# Classification and clustering of Guediawaye census data

## Geodemographic classification

```{r}
set.seed(1994)

dfw_kmeans_2013 <- dfw_pca_2013 %>%
  st_drop_geometry() %>%
  select(PC1:PC8) %>%
  kmeans(centers = 6)

table(dfw_kmeans_2013$cluster)

##
dfw_kmeans_2002 <- dfw_pca_2002 %>%
  st_drop_geometry() %>%
  select(PC1:PC8) %>%
  kmeans(centers = 6)

table(dfw_kmeans_2002$cluster)
```

```{r}
dfw_clusters_2013 <- dfw_pca_2013 %>%
  mutate(cluster = as.character(dfw_kmeans_2013$cluster))

ggplot(dfw_clusters_2013, aes(fill = cluster)) + 
  geom_sf(size = 0.1) + 
  scale_fill_brewer(palette = "Set1") + 
  labs(title = "Map of geodemographic clusters for Guediawaye Census 2013") +
  theme_void() + 
  labs(fill = "Cluster ")
```


```{r}
dfw_clusters_2002 <- dfw_pca_2002 %>%
  mutate(cluster = as.character(dfw_kmeans_2002$cluster))

ggplot(dfw_clusters_2002, aes(fill = cluster)) + 
  geom_sf(size = 0.1) + 
  scale_fill_brewer(palette = "Set1") + 
  labs(title = "Map of geodemographic clusters for Guediawaye Census 2002") +
  theme_void() + 
  labs(fill = "Cluster ")
```
```{r}
dfw_clusters <- dfw_clusters_2002 %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(dfw_clusters_2013 %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


dfw_clusters %>% 
    
      ggplot(aes(fill = cluster)) + 
      
  geom_sf(size = 0.1) + 
  scale_fill_brewer(palette = "Set1") + 
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            panel.grid = element_line(color = "white", size = 0.8))
```

```{r}
library(plotly)

cluster_plot <- ggplot(dfw_clusters_2013, 
                       aes(x = PC1, y = PC2, color = cluster)) + 
  geom_point() + 
  scale_color_brewer(palette = "Set1") + 
  labs(title = "Interactive scatterplot of PC1 and PC2 colored by cluster for Guediawaye Census 2013") +
  theme_minimal()

ggplotly(cluster_plot) %>%
  layout(legend = list(orientation = "h", y = -0.15, 
                       x = 0.2, title = "Cluster"))
```


```{r}
cluster_plot <- ggplot(dfw_clusters_2002, 
                       aes(x = PC1, y = PC2, color = cluster)) + 
  geom_point() + 
  scale_color_brewer(palette = "Set1") + 
  labs(title = "Interactive scatterplot of PC1 and PC2 colored by cluster for Guediawaye Census 2002") +
  theme_minimal()

ggplotly(cluster_plot) %>%
  layout(legend = list(orientation = "h", y = -0.15, 
                       x = 0.2, title = "Cluster"))
```


## Spatial clustering & regionalization

```{r}
library(spdep)

input_vars <- dfw_pca_2013 %>%
  select(PC1:PC8) %>%
  st_drop_geometry() %>%
  as.data.frame() 

skater_nbrs <- poly2nb(dfw_pca_2013, queen = TRUE)
costs <- nbcosts(skater_nbrs, input_vars)
skater_weights <- nb2listw(skater_nbrs, costs, style = "B")
```

```{r}
mst <- mstree(skater_weights)

regions <- skater(
  mst[,1:2], 
  input_vars, 
  ncuts = 7,
  crit = 10
)
```


```{r}
dfw_clusters_2013$region <- as.character(regions$group)

ggplot(dfw_clusters_2013, aes(fill = region)) + 
  geom_sf(size = 0.1) + 
  labs(title = "Map of contiguous regions derived with the SKATER algorithm for Guediawaye Census 2013") +
  scale_fill_brewer(palette = "Set1") + 
  theme_void()
```
```{r}
input_vars <- dfw_pca_2002 %>%
  select(PC1:PC8) %>%
  st_drop_geometry() %>%
  as.data.frame() 

skater_nbrs <- poly2nb(dfw_pca_2002, queen = TRUE)
costs <- nbcosts(skater_nbrs, input_vars)
skater_weights <- nb2listw(skater_nbrs, costs, style = "B")
```

```{r}
mst <- mstree(skater_weights)

regions <- skater(
  mst[,1:2], 
  input_vars, 
  ncuts = 7,
  crit = 10
)
```


```{r}
dfw_clusters_2002$region <- as.character(regions$group)

ggplot(dfw_clusters_2002, aes(fill = region)) + 
  geom_sf(size = 0.1) + 
  labs(title = "Map of contiguous regions derived with the SKATER algorithm for Guediawaye Census 2002") +
  scale_fill_brewer(palette = "Set1") + 
  theme_void()
```


```{r}
dfw_clusters <- dfw_clusters_2002 %>%
  dplyr::mutate(RGPH = "Census 2002") %>% 
  plyr::rbind.fill(dfw_clusters_2013 %>%
  dplyr::mutate(RGPH = "Census 2013")) %>% 
  st_as_sf()


dfw_clusters %>% 
    
      ggplot(aes(fill = region)) + 
      
  geom_sf(size = 0.1) + 
  scale_fill_brewer(palette = "Set2") + 
      facet_wrap(~RGPH) +
      
      # Adjusting the plot theme
      theme_map(base_size = 8) +
      theme(panel.background = element_rect(),
            #legend.background = element_blank(),
            axis.ticks = element_blank(),
            axis.text = element_blank(),
            legend.position = "bottom", 
            text = element_text(size = 8), 
            #panel.grid = element_line(color = "white", size = 0.8),
            legend.box.background = element_rect(),
            legend.box.margin = margin(6, 6, 6, 6),
            plot.background = element_rect(fill = "white"))+
  ggspatial::annotation_scale(
    location = "br",
    bar_cols = c("grey60", "white"),
    text_family = "ArcherPro Book"
  ) +
  ggspatial::annotation_north_arrow(
    location = "tl", which_north = "true",
    #pad_x = unit(0.4, "in"), pad_y = unit(0.4, "in"),
    #which_north = "grid",
  height = unit(1, "cm"),
  width = unit(1, "cm"),
  )
```


