---
title: "Cleaning File"
author: "IBRAHIM KASSOUM Habibou"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, error=FALSE)
```


## Project Sénégal

### Preambule: 

```{r Package needed}

## Importing library
### List of required packages
required_packages <- c("tidyverse","janitor" ,"readr","dplyr","haven","sf","sp")

# Check if packages are installed
missing_packages <- setdiff(required_packages, installed.packages()[,"Package"])

### Install missing packages
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

### Load all packages
lapply(required_packages, library, character.only = TRUE)

```


### Importing the dataset

```{r}

# Read shapefile data for 2002 and 2013
sp_rgph_2002 <- sf::read_sf(paste0(here::here(),"/data/Données/SHAPEFILE/EAs_2002.shp"))

sp_rgph_2013 <- sf::read_sf(paste0(here::here(),"/data/Données/SHAPEFILE/EAs_2013.shp"))

# Display the first 10 rows of the 2002 and 2013 shapefiles
head(sp_rgph_2002, 10L)
head(sp_rgph_2013, 10L)
```
### Importing the dataset an checking for duplicated row

```{r}
# Read and check for duplicated rows in 2002 datasets

deces_2002 <- read_sav(paste0(here::here(),"/data/Données/1. RGPH 2002/Requete PDBA2k2 Deces.sav"))
deces_2002 %>% janitor::get_dupes() 


individus_2002 <- haven::read_sav(paste0(here::here(),"/data/Données/1. RGPH 2002/Requete PDBA2k2 Individus.v2.sav"))
individus_2002 %>% janitor::get_dupes() 


menage_2002 <- haven::read_sav(paste0(here::here(),"/data/Données/1. RGPH 2002/Requete PDBA2k2 menage.sav"))
menage_2002 %>% janitor::get_dupes() 
```


```{r}

# Read and check for duplicated rows in 2013 datasets
deces_2013 <- read_sav(paste0(here::here(),"/data/Données/4. RGPH 2013/Requete PDBA2k13 Deces_Guediawaye.sav"))
deces_2013 %>% janitor::get_dupes() 


individus_2013 <- haven::read_sav(paste0(here::here(),"/data/Données/4. RGPH 2013/Requete PDBA2k13 Indidivus GUEDIAWAYE.sav"))
individus_2013 %>% janitor::get_dupes() 


menage_2013 <- haven::read_sav(paste0(here::here(),"/data/Données/4. RGPH 2013/Requete PDBA2k13 menage.sav"))
menage_2013 %>% janitor::get_dupes() 
```


```{r}

# 2013
lst_dr_sp <- sp_rgph_2013 %>% as.data.frame() %>% select(COD_DR2012) %>% pull()  

lst_dr_menage <- menage_2013 %>% select(IDDR) %>% pull() 

```


### -	Calcul des indices sur les variables composants liés au ménage (Age des chefs de ménage, par sexe entre les deux recensements

```{r}
# Select the CM and compute it average age
individus_2013 %>% 
  
  # Transform all "haven_labelled" to factor variable
  #labelled::unlabelled() %>% 
  
  ##  Select Household head informations
  filter(B04==1) %>% 
  
  # Grouping by sexe
  group_by(B06) %>% 
  
  # Computing number of HH and average age
  summarise("Nombre d'individus" = n(), 
            "Age moyen" = mean(B08, na.rm = TRUE))
```



### -	Calcul multivarié pour des indices de pauvreté (bien des ménages, Niveau d’instruction, accès aux services (bien d’aisance, eau, etc.)


### -	Calcul indice sur la dynamique de la population (croissance démographique au niveau des DR)



